---
title: 'Baby mental life: Bi-cross-validation'
author: "Kara Weisman"
date: "2025-02-04"
output: html_document
---

```{r setup, set.seed(94305), include = F}
knitr::opts_chunk$set(cache = T)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(fig.pos = 'H')
knitr::opts_chunk$set(include = F)
```

```{r load packages}
# library(ggdendro)
# library(dendextend)
library(psych)
library(lme4)
library(mgcv)
library(kableExtra)
library(langcog)
library(parameters)
library(cowplot)
# library(fda)
library(rsq)
library(ggnewscale)
library(tidyverse)

library(softImpute)

conflicted::conflict_prefer("select", "dplyr")
conflicted::conflict_prefer("filter", "dplyr")
```

```{r load custom funs}
source("../code/custom_funs.R")
```

```{r load data}
# suppress warnings and messages because dplyr is verbose
suppressWarnings(suppressMessages(source("../code/data_prep_s1.R", chdir = TRUE)))
suppressWarnings(suppressMessages(source("../code/data_prep_s2.R", chdir = TRUE)))
suppressWarnings(suppressMessages(source("../code/data_prep_s3.R", chdir = TRUE)))
```


\newpage
In a series of large-scale studies we assessed how US adults conceptualize the development of the human mind over the first five years of life. We identified four core capacities that are perceived by US adults to develop over human infancy and early childhood, _bodily sensation_, _negative affect_, _social connection_, and _cognition and control_ (Study 1); charted how these different aspects of mental life are perceived to change over the first five years of a child's life (Studies 2-3); and explored the intuitive theories that underlie the observed differences in the perceived developmental trajectories of these four aspects of mental life (Study 3). 

In this RMarkdown file, we apply a "bi-cross-validation" approach to the EFA in Study 1.


```{r load custom funs}
source("../code/custom_funs.R")
```

```{r load data}
# suppress warnings and messages because dplyr is verbose
suppressWarnings(suppressMessages(source("../code/data_prep_s1.R", chdir = TRUE)))
# suppressWarnings(suppressMessages(source("../code/data_prep_s2.R", chdir = TRUE)))
# suppressWarnings(suppressMessages(source("../code/data_prep_s3.R", chdir = TRUE)))
```


### BCV

```{r}
# had to instsall from source, not maintained on CRAN. retrieved from https://jingshuw.org/publication/owen-2016-aa/ (see also https://cran.r-project.org/web/packages/esaBcv/)
# install.packages("https://cran.r-project.org/src/contrib/Archive/esaBcv/esaBcv_1.2.1.1.tar.gz")

# library(esaBcv)
```

```{r}
# X <- as.matrix(d1_all)
# 
# bcv_res <- EsaBcv(X, center = F) # center adds intercept to model
# 
# plot(bcv_res)
```

```{r}
# bcv_res$best.r
# bcv_res$mu
```

```{r}
# bcv_res_esa <- ESA(Y = d1_all, r = bcv_res$best.r)
# 
# bcv_res_esa_V <- bcv_res_esa$estV
# colnames(bcv_res_esa_V) <- paste0("V", 1:bcv_res$best.r) 
# rownames(bcv_res_esa_V) <- colnames(d1_all)
# 
# fa.sort(bcv_res_esa_V)
```

```{r}
# bcv_res_efa <- fa(d1_all, nfactors = bcv_res$best.r, rotate = "oblimin")
```

```{r, fig.width = 10, fig.asp = 1.1}
# heatmap_fun(bcv_res_efa)
```



```{r}
# library(psych)
# library(dplyr)
# library(tidyr)

# Example Data: Assuming d1_all includes ParticipantID
X <- as.matrix(d1_all)  # Data matrix (participants x items)

# Step 1: Create Cross-Validation folds (leave out entire participants)
k_folds_row <- 5  # Number of participant folds
k_folds_col <- 5  # Number of item folds
set.seed(94025)  # For reproducibility

# Assign participants to folds_row
unique_participants <- unique(d1$ResponseId)

participant_folds <- sample(rep(1:k_folds_row, length.out = length(unique_participants)))
names(participant_folds) <- unique_participants

participant_folds_target00mo <- participant_folds
names(participant_folds_target00mo) <- paste(unique_participants, "target00mo", sep = "_")

participant_folds_target09mo <- participant_folds
names(participant_folds_target09mo) <- paste(unique_participants, "target09mo", sep = "_")

participant_folds_target60mo <- participant_folds
names(participant_folds_target60mo) <- paste(unique_participants, "target60mo", sep = "_")

participant_folds <- c(
  participant_folds_target00mo,
  participant_folds_target09mo,
  participant_folds_target60mo)

# Assign items (columns) to folds_col
unique_items <- colnames(X)

item_folds <- sample(rep(1:k_folds_col, length.out = length(unique_items)))
names(item_folds) <- unique_items

# Map fold assignments back to rows and columns
folds_row <- participant_folds[as.character(rownames(X))]  # Assign rows to folds based on participant ID
folds_col <- item_folds[as.character(colnames(X))]  # Assign columns to folds

# Step 2: Bi-Cross-Validation Function for Varying Number of Factors
nfact_range <- 1:12  # Test from 1 to 12 factors

# Initialize storage for results
bcv_results <- lapply(nfact_range, function(nfact) {
  fold_metrics_list <- lapply(1:k_folds_row, function(fold_row) {
    fold_metrics_col_list <- lapply(1:k_folds_col, function(fold_col) {
      
      # Training and Test Data
      train_data <- X
      test_data <- X[folds_row == fold_row, , drop = FALSE]  # Hold out participants (rows)
      
      train_data[folds_row == fold_row, ] <- NA  # Mask participant data
      train_data[, folds_col == fold_col] <- NA  # Mask item (column) data
      
      # Step 3: Remove items with excessive missing data
      missing_prop <- colMeans(is.na(train_data))
      items_to_keep <- names(missing_prop[missing_prop < 0.5])  # Keep items with <50% missingness
      train_data_filtered <- train_data[, items_to_keep, drop = FALSE]
      test_data_filtered <- test_data[, items_to_keep, drop = FALSE]
      
      # Check if enough items are left to perform EFA
      if (ncol(train_data_filtered) < nfact) return(rep(NA, 5))  # Return NA for all metrics
      
      # Step 4: Fit EFA on Training Data
      efa_fit <- tryCatch({
        fa(train_data_filtered, nfactors = nfact, 
           rotate = "oblimin", fm = "minres", missing = TRUE, scores = "tenBerge")
      }, error = function(e) { return(NULL) })
      
      # Skip fold if EFA did not converge
      if (is.null(efa_fit)) return(rep(NA, 5))
      
      # Step 5: Predict Factor Scores for Test Data Using Observed Values
      predicted_factor_scores <- tryCatch({
        factor.scores(test_data_filtered, efa_fit, method = "tenBerge")$scores
      }, error = function(e) { return(NULL) })
      
      # Skip fold if factor score prediction fails
      if (is.null(predicted_factor_scores)) return(rep(NA, 5))
      
      # Step 6: Reconstruct Test Data Using Predicted Factor Scores
      reconstructed_data <- predicted_factor_scores %*% t(efa_fit$loadings[])
      
      # Align reconstructed data with original test_data columns
      reconstructed_full <- matrix(NA, nrow = nrow(test_data), ncol = ncol(X))
      colnames(reconstructed_full) <- colnames(X)
      reconstructed_full[, items_to_keep] <- reconstructed_data
      
      # Identify positions with observed data in both matrices
      mask <- !is.na(test_data) & !is.na(reconstructed_full)
      
      # Step 7: Compute Multiple Evaluation Metrics
      if (sum(mask) > 0) {  
        mse <- mean((test_data[mask] - reconstructed_full[mask])^2)
        rmse <- sqrt(mse)
        mae <- mean(abs(test_data[mask] - reconstructed_full[mask]), na.rm = TRUE)
        srmr <- sqrt(mean((cor(test_data, use = "pairwise.complete.obs") - 
                           cor(reconstructed_full, use = "pairwise.complete.obs"))^2, na.rm = TRUE))
        correlation <- cor(as.vector(test_data[mask]), as.vector(reconstructed_full[mask]), use = "pairwise.complete.obs")
      } else {
        mse <- NA
        rmse <- NA
        mae <- NA
        srmr <- NA
        correlation <- NA
      }
      
      return(c(MSE = mse, RMSE = rmse, MAE = mae, SRMR = srmr, Correlation = correlation))
      
    })  # End column fold loop
    
    # Convert fold-level metrics to a data frame and take the mean across item folds
    fold_metrics_col_df <- do.call(rbind, fold_metrics_col_list)
    colMeans(fold_metrics_col_df, na.rm = TRUE)
    
  })  # End row fold loop
  
  # Convert row-level metrics to a data frame and take the mean across participant folds
  fold_metrics_df <- do.call(rbind, fold_metrics_list)
  colMeans(fold_metrics_df, na.rm = TRUE)
})

# Convert list to data frame with explicit column names
bcv_results_df <- as.data.frame(do.call(rbind, bcv_results)) %>%
  rownames_to_column("nfact") %>%
  mutate(nfact = as.numeric(nfact)) %>%
  select(nfact, everything())

colnames(bcv_results_df)[2:ncol(bcv_results_df)] <- c("MSE", "RMSE", "MAE", "SRMR", "Correlation")

# Display the full table of evaluation metrics
bcv_results_df %>%
  kable(digits = 2) %>%
  kable_styling()
```

```{r}
# library(ggplot2)
# library(dplyr)
# library(tidyr)

for(i in colnames(bcv_results_df)[2:ncol(bcv_results_df)]) {
  which_metric <- i
  
  # Compute first and second derivatives to detect the elbow point
  first_diff <- diff(bcv_results_df[[which_metric]])  # Use [[ ]] to subset correctly
  second_diff <- diff(first_diff)
  
  # Find the first local minimum of the second derivative
  if (i == "Correlation") {
    elbow_point <- which(diff(sign(diff(second_diff))) < 0)[1] + 1
  } else {
    elbow_point <- which(diff(sign(diff(second_diff))) > 0)[1] + 1
  }
  
  # Plot with the elbow point highlighted
  p <- bcv_results_df %>%
    ggplot(aes(x = nfact, y = .data[[which_metric]])) +  # Use .data[[ ]] for tidyverse
    geom_line() +
    geom_point() +
    scale_x_continuous("Number of factors", breaks = seq(0, 100, 1)) +
    annotate("point", x = elbow_point, y = bcv_results_df[[which_metric]][elbow_point],
             color = "red", shape = 4, size = 10) +
    theme_bw()
  
  print(p)
}
```

```{r, fig.width = 10, fig.asp = 1}
fa(d1_all, nfactors = 4, rotate = "geominQ") %>% heatmap_fun()
fa(d1_all, nfactors = 5, rotate = "geominQ") %>% heatmap_fun()
fa(d1_all, nfactors = 6, rotate = "geominQ") %>% heatmap_fun()
```

```{r, fig.width = 10, fig.asp = 1}
fa(d1_all, nfactors = 4) %>% loadings(cutoff = 0) %>% fa.sort()
fa(d1_all, nfactors = 5) %>% loadings(cutoff = 0) %>% fa.sort()
fa(d1_all, nfactors = 6) %>% loadings(cutoff = 0) %>% fa.sort()
```
