---
# title: "Beliefs about the development of mental life"
# subtitle: "Supplemental Materials"
# author: "Kara Weisman, Lucy S. King, and Kathryn Humphreys"
# date: "2024-12-30"
output:
  bookdown::pdf_document2:
    includes:
      in_header: header.tex
      before_body: cover.tex
    latex_engine: xelatex
    fig_caption: yes
    toc: true
    toc_depth: 2
    number_sections: false
  #   reference_docx: word-styles-ref.docx
  # word_document:
bibliography: "references.bib"
csl: apa.csl
# cls: article.cls
figsintext: yes
header-includes: 
   - \usepackage{lineno}
   - \linenumbers
   - \modulolinenumbers[5]
   - \usepackage{float}
   - \usepackage{pdflscape}
tables: yes
---

```{r setup, set.seed(94305), include = F}
knitr::opts_chunk$set(cache = T)
knitr::opts_chunk$set(echo = F)
knitr::opts_chunk$set(fig.pos = 'H')
knitr::opts_chunk$set(include = F)
```

```{r resave plots}
# do you want to resave the plots?
resave_plots <- T # uncomment if yes
# resave_plots <- F # uncomment if no
```

```{r load packages}
library(ggdendro)
library(dendextend)
library(psych)
library(lme4)
library(mgcv)
library(kableExtra)
library(langcog)
library(parameters)
library(cowplot)
library(tidyverse)
library(rsq)
library(ggnewscale)
```

```{r load custom funs}
source("../code/custom_funs.R")
```

```{r load data}
# suppress warnings and messages because dplyr is verbose
suppressWarnings(suppressMessages(source("../code/data_prep_s1.R", chdir = TRUE)))
suppressWarnings(suppressMessages(source("../code/data_prep_s2.R", chdir = TRUE)))
suppressWarnings(suppressMessages(source("../code/data_prep_s3.R", chdir = TRUE)))
```


\newpage
In a series of large-scale studies we assessed how US adults conceptualize the development of the human mind over the first five years of life. We identified four core capacities that are perceived by US adults to develop over human infancy and early childhood, _bodily sensation_, _negative affect_, _social connection_, and _cognition and control_ (Study 1); charted how these different aspects of mental life are perceived to change over the first five years of a child's life (Studies 2-3); and explored the intuitive theories that underlie the observed differences in the perceived developmental trajectories of these four aspects of mental life (Study 3). 

In these Supplemental Materials, we provide extended information about the methods and results presented in the main paper, present additional results, and provide an item-by-item comparison of the capacity items used here to items used in previous work on mind perception [@gray.etal2007; @weisman.etal2017a; @weisman.etal2017b; @malle2019].

Note: This is a fully reproducible manuscript, written in RMarkdown using the "knitr" package for R [@xie2021]. The full, editable script for all analyses, figures, and tables (here and in the main text) is available in the .Rmd file that generated this document, which can be found in the following GitHub repository: https://github.com/kgweisman/baby_mental_life_ms.


# Extended methods

```{r}
# NOTE: this is a bit of a circular process -- results files are generated in this .Rmd file, but they also need to be loaded in order to write the Extended Methods section. For this reason, results RDS files are included in the github repo.
# To re-make results from scratch, comment out these lines, run this whole .Rmd file, and then uncomment these lines before knitting the supplement.

s1_results <- readRDS("../outputs/s1_results.rds")
s2_results <- readRDS("../outputs/s2_results.rds")
s3_results <- readRDS("../outputs/s3_results.rds")

# ALSO NOTE: if you encounter problems knitting, try clearing the knit cache.
```

Methods, inclusion/exclusion criteria, and analyses marked as "preregistered" were preregistered on the Open Science Framework (OSF) website (Study 1: https://osf.io/e6ajh; Study 2: https://osf.io/j72dg; Study 3: https://osf.io/xh8ce). 

All materials used in these studies are available in the following OSF repository: https://osf.io/xrznd/ (DOI: 10.17605/OSF.IO/XRZND).

All studies were approved by the Institution Review Board for Human Subjects at Stanford University. 


## Study 1

### Participants

`r s1_results$demo %>% distinct() %>% nrow()` US adults participated via Amazon Mechanical Turk (MTurk) in July–August 2018. To be eligible for the study, participants were required to have gained approval for at least 95% of their previous work on MTurk, with at least 50 HITs completed; have MTurk accounts based in the US; indicate that they were between the ages of 18-45 years; and report that their English abilities were either "advanced" or "superior" (not "novice," "intermediate," or decline to provide information about English ability). Participants were paid $2.00 for approximately 10-15 minutes of their time.

According to self-report, participants ranged in age from `r min(s1_results$demo$Age)`-`r max(s1_results$demo$Age)` years ($M$ = `r round_x(mean(s1_results$demo$Age, na.rm = T), 2)` years, $sd$ = `r round_x(sd(s1_results$demo$Age, na.rm = T), 2)` years) and included more men (`r round((s1_results$demo %>% count(GenderSex) %>% mutate(prop = n/sum(n)) %>% filter(GenderSex == "Male") %>% pull(prop))*100, 0)`%) than women (`r round((s1_results$demo %>% count(GenderSex) %>% mutate(prop = n/sum(n)) %>% filter(GenderSex == "Female") %>% pull(prop))*100, 0)`%; <1% of participants identified as some other gender or declined to disclose). Participants predominantly identified as White (`r round((s1_results$demo %>% count(RaceEthnicity) %>% mutate(prop = n/sum(n)) %>% filter(RaceEthnicity == "White") %>% pull(prop))*100, 0)`%; <`r round(100*(s1_results$demo %>% count(RaceEthnicity) %>% mutate(prop = n/sum(n)) %>% filter(RaceEthnicity != "White") %>% filter(prop == max(prop)) %>% pull(prop)) + 1, 0)`% identified as any other race/ethnicity, identified as more than one race/ethnicity, or declined to disclose). `r round(100*(s1_results$demo %>% count(Education) %>% arrange(desc(Education)) %>% mutate(cumprop = cumsum(n)/sum(n)) %>% filter(grepl("Bachelor", Education)) %>% pull(cumprop)))`% of participants had obtained at least a Bachelor's degree. `r round(100*(s1_results$demo %>% count(Parent) %>% mutate(prop = n/sum(n)) %>% filter(Parent == "Yes") %>% pull(prop)))`% of participants indicated that they were parents.

An additional 77 participants completed the study but were excluded for failing one or more (out of three) embedded attention checks (e.g., failing to choose "34" in response to the question "Please select 34") or for indicating via a multiple choice question at the end of the survey that they had not paid attention, had not avoided distractions, or had not taken the study tasks seriously.

### Materials and procedure

After completing a screening questionnaire (in which participants provided information about their gender/sex, age, and English abilities) and providing informed consent, participants were provided with an overview of the study, as follows: "In this study, we will ask you about the mental capacities of children of different ages, from newborns to 5-year-olds. We will show you example photos of children at three of the following ages to remind you of what children look like at that age, but please think about children of that age in general when answering these questions. There are 60 brief questions for each of the three ages (for example, 'To what extent is a newborn capable of feeling happy?'), followed by a few demographic questions about yourself. We are interested in your opinions about these questions. Please do not look up extra information about these questions—we just want to know what you think." Participants were shown an array of pictures of children at several target ages between 0-5 years; see "Target ages and accompanying photographs," below.

Participants then proceeded through three trials in which they assessed the mental lives of children at three target ages: birth, 9 months, and 5 years. On each trial, they were shown two representative photographs of children at the target age (with the photograph labeled "newborns," "9-month-olds," or "5-year-olds"), and asked to answer the following question for 60 capacities: "To what extent is a [newborn/9-month-old/5-year-old] capable of [this capacity]?" Participants responded on a sliding scale from 0 (labeled as "not at all capable") to 100 (labeled as "completely capable").

On the final page of the survey, participants were asked to provide additional demographic information.

#### Capacities

The 60 capacities included in Study 1 were drawn from several areas of previous research. Our goals in selecting items were to: (1) balance the representation of the conceptual organization underlying mental capacity attributions in general, as identified in previous work [@gray.etal2007; @weisman.etal2017a; @weisman.etal2017b; @weisman.etal2018; @malle2019]; (2) add items relevant to early development, as identified by existing measures of temperament and behavior in early life [@putnam.etal2006; @rothbart1978]; and (3) assess the broadest range of mental capacities within financial and participant burden constraints, as identified by extensive discussions among the authors drawing on our own research and clinical experience with infants and young children. 

This process yielded a list of 60 capacities. Roughly half of this list (29 of 60 items) were drawn from (or closely based on) the items included in Weisman et al.'s recent studies [@weisman.etal2017a]: "being aware of things", "feeling calm", "feeling embarrassed", "feeling guilty", "feeling happy", "feeling love", "feeling pain", "feeling pleasure", "feeling pride", "feeling sad", "feeling safe", "feeling scared", "feeling tired", "getting angry", "getting hungry", "getting hurt feelings", "having goals", "having self-control", "having thoughts", "having wants and desires", "hearing sounds", "making choices", "planning", "reasoning about things", "recognizing others' emotions", "recognizing somebody else", "remembering things", "seeing", and "telling right from wrong". By extension, the list included close variants on 14 of the 20 capacities evaluated in Gray et al.'s original study [@gray.etal2007]. 

The other half (31 of 60 items) were novel items, not used in previous studies of mind perception, but common in clinical assessments of infant temperament and behavior: "being afraid of somebody", "being angry at somebody", "being comforted by physical touch", "calming themselves down", "controlling their emotions", "detecting danger", "feeling annoyed", "feeling bored", "feeling confused", "feeling distressed", "feeling excited", "feeling frustrated", "feeling helpless", "feeling hopeless", "feeling lonely", "feeling neglected", "feeling overwhelmed", "feeling physically uncomfortable", "feeling textures (for example, 'smooth', 'rough')", "feeling thirsty", "feeling too hot or too cold", "feeling worried", "finding something funny", "focusing on a goal", "getting pleasure from music", "imagining things", "learning from other people", "listening to somebody", "loving somebody", "thinking before they act", and "understanding what somebody else is thinking".

See "Selection of capacities: Comparison to previous studies" for an item-by-item comparison of the capacities used here to those used in previous studies of mind perception [@gray.etal2007; @weisman.etal2017a; @weisman.etal2017b; @malle2019].

Capacities were presented in a random order each time they were displayed (i.e., order varied across both target ages and participants).

#### Target ages and accompanying photographs

In planning this series of studies, we selected 13 target ages based on the developmental milestone ages identified by the CDC [@CDC2021] with the addition of the American Academy of Pediatrics' recommendations for pediatrician visits before 2 months [@AAP2018]: birth, 4 days, 1 month, 2 months, 4 months, 6 months, 9 months, 12 months, 18 months, 2 years, 3 years, 4 years, and 5 years. We limited Study 1 to just the youngest, median, and oldest target ages—presented to participants as "newborns," "9-month-olds" and "5-year-olds"—in order to maximize the number of capacities included in the study without over-burdening participants. 

In selecting photos for each target age, we aimed to have photos of the same (or very similar-looking) children at all target ages so as to emphasize the question of the development of mental capacities over the course of an individual child's early life. We chose to illustrate each target age with two photos in order to emphasize that we were interested in children's mental capacities in general, rather than the experiences of a particular child in a particular moment. We also presented participants with instructions that underscored this point (see above). We included one boy and one girl at each target age in order to generalize across perceived sex/gender differences in children's early mental life. We used photographs of two (apparently) White, non-Hispanic children, for three reasons: (1) In order to avoid introducing spurious differences across target ages by confounding a target's age and race/ethnicity; (2) In order to match the anticipated race/ethnicity of the majority of participants on MTurk; and (3) Because of the limited set of high-resolution photographs of individual children of color at different points of development currently available to us. Conducting similar studies of adults' perceptions of the development of mental life with more ethnically and racially diverse participants, and examining the effects of sex, gender, race, and ethnicity on adults' perceptions of children's mental life, will be critical next steps in this line of research.

Target ages were presented in chronological order, with participants first assessing all 60 capacities for newborns, then 9-month-old infants, and finally 5-year-old children.

### Analysis plan

To our knowledge, there is currently no consensus about how to account for within-participant variance alongside between-participant variance in the context of exploratory factor analysis or similar dimensionality reduction analyses. Thus, to conduct the EFAs reported in Results, we proceeded as if each participant's ratings of the 60 mental capacities for each of the three target ages were a distinct source of data; in other words, as if we had 903 participants assessing one target age each, instead of 301 participants assessing three target ages each. (See [@weisman.etal2017a; @weisman.etal2021], for similar treatments of within-subjects data in the context of EFA.) We used Pearson correlations to obtain minimal residual solutions and applied an oblique transformation (oblimin) to the resulting factor structures, using the "psych" package for R [@revelle2021]. 

In order to determine how many factors to retain, we examined the results of three factor retention protocols: (1) Parallel analysis, which compares the observed correlation structure to the correlation structure arising from random datasets of the same size; (2) Minimizing the Bayesian Information Criterion (BIC), which is one method of optimizing both goodness of fit and parsimony; and (3) A set of factor retention criteria that have been used in Weisman et al.'s previous work [@weisman.etal2017a], in which they retained factors with eigenvalues greater than 1.0, which individually accounted for greater than 5% of the shared variance before transformation or rotation, and which were the "dominant" factor (the factor with the strongest absolute factor loading) for at least one mental capacity after transformation (a factor had to meet all three of these criteria in order to be retained). Our interpretation of how best to characterize our dataset (i.e., how many factors we observe) was guided by the degree of consensus among these three protocols and the interpretability of the retained factors under each protocol.

Parallel analysis suggested a `r s1_results$efa$factors`-factor solution (accounting for `r round(100*(s1_results$efa$Vaccounted["Cumulative Var",ncol(s1_results$efa$Vaccounted)]))`% of the total variance); minimizing BIC suggested a `r s1_results$efa_bic$factors`-factor solution (accounting for `r round(100*(s1_results$efa_bic$Vaccounted["Cumulative Var",ncol(s1_results$efa_bic$Vaccounted)]))`% of the total variance); and Weisman et al.'s (2017) factor retention criteria suggested a `r s1_results$efa_wdm$factors`-factor solution (accounting for `r round(100*(s1_results$efa_wdm$Vaccounted["Cumulative Var",ncol(s1_results$efa_wdm$Vaccounted)]))`% of the total variance). We present the `r s1_results$efa$factors`-factor solution in the main text because it was of intermediate size and because we judged the factors it revealed to be intelligible and of theoretical interest; we present all three solutions in these Supplemental Materials.

## Study 2

### Participants

`r s2_results$demo %>% distinct() %>% nrow()` US adults participated via MTurk in August 2018. Eligibility requirements were identical to Study 1. Participants were paid $2.90 for approximately 15-20 minutes of their time. In response to the overrepresentation of men in Study 1, in Study 2 we recruited men and women through separate "Human Intelligence Tasks" (HITs).

According to self-report, participants ranged in age from `r min(s2_results$demo$Age)`-`r max(s2_results$demo$Age)` years ($M$ = `r round_x(mean(s2_results$demo$Age, na.rm = T), 2)` years, $sd$ = `r round_x(sd(s2_results$demo$Age, na.rm = T), 2)` years) and (per our recruitment strategy) included roughly equal numbers of men (`r round((s2_results$demo %>% count(GenderSex) %>% mutate(prop = n/sum(n)) %>% filter(GenderSex == "Male") %>% pull(prop))*100, 0)`%) and women (`r round((s2_results$demo %>% count(GenderSex) %>% mutate(prop = n/sum(n)) %>% filter(GenderSex == "Female") %>% pull(prop))*100, 0)`%). Participants predominantly identified as White (`r round((s2_results$demo %>% count(RaceEthnicity) %>% mutate(prop = n/sum(n)) %>% filter(RaceEthnicity == "White") %>% pull(prop))*100, 0)`%; <`r round(100*(s2_results$demo %>% count(RaceEthnicity) %>% mutate(prop = n/sum(n)) %>% filter(RaceEthnicity != "White") %>% filter(prop == max(prop)) %>% pull(prop)) + 1, 0)`% identified as any other race/ethnicity, identified as more than one race/ethnicity, or declined to disclose). `r round(100*(s2_results$demo %>% count(Education) %>% arrange(desc(Education)) %>% mutate(cumprop = cumsum(n)/sum(n)) %>% filter(grepl("Bachelor", Education)) %>% pull(cumprop)))`% of participants had obtained at least a Bachelor's degree. `r round(100*(s2_results$demo %>% count(Parent) %>% mutate(prop = n/sum(n)) %>% filter(Parent == "Yes") %>% pull(prop)))`% of participants indicated that they were parents.

An additional 90 participants completed the study but were excluded for failing one or more (out of four) attention checks or indicating that they had not paid attention (see Study 1).

### Materials and procedure

After completing a screening questionnaire (in which participants provided information about their gender/sex, age, and English abilities) and providing informed consent, participants were provided with an overview of the study that was nearly identical to that used in Study 1.

Participants then proceeded through 13 trials in which they assessed the mental lives of children at 13 target ages: birth, 4 days, 1 month, 2 months, 4 months, 6 months, 9 months, 12 months, 18 months, 2 years, 3 years, 4 years, and 5 years. On each trial, they were shown two representative photographs of children at the target age, and asked to answer the following question for 20 capacities: "To what extent is a [newborn/4-day-old/etc.] capable of [this capacity]?" As in Study 1, participants responded on a sliding scale from 0 (labeled as "not at all capable") to 100 (labeled as "completely capable").

On the final page of the survey, participants were asked to provide additional demographic information.

#### Capacities

In order to ask participants to assess a more fine-grained array of target ages without undue participant burden, we limited our list of capacities to 20 of the 60 capacities used in Study 1. To identify this list, we drew on the results of Study 1, selecting 5 items for each of the 4 factors identified there. For each factor, we aimed to select items that loaded strongly on that factor, did not cross-load strongly on other factors, were sufficiently distinguishable from each other in meaning (i.e., not overly redundant), and captured our qualitative understanding of the latent construct that each factor corresponded to. 

To represent _bodily sensation_, we chose the following items: "getting hungry", "feeling pain", "feeling tired", "feeling physically uncomfortable", and "hearing sounds". In the oblimin-transformed `r s1_results$efa$factors`-factor solution from Study 1 all of these items had loadings $\geq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("getting_hungry", "feeling_pain", "feeling_tired", "feeling_physically_uncomfortable", "hearing_sounds")) %>% pull("MR2") %>% min() %>% round_x(., 2)` on this factor, and absolute loadings $\leq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("getting_hungry", "feeling_pain", "feeling_tired", "feeling_physically_uncomfortable", "hearing_sounds")) %>% gather(factor, loading, -capacity) %>% filter(factor != "MR2") %>% pull(loading) %>% max((abs(.))) %>% round_x(., 2)` on the other three factors.

To represent _negative affect_, we chose the following items: "feeling distressed", "feeling lonely", "feeling frustrated", "feeling helpless", and "feeling overwhelmed". In Study 1 all of these items had loadings $\geq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("feeling_distressed", "feeling_lonely", "feeling_frustrated", "feeling_helpless", "feeling_overwhelmed")) %>% pull("MR4") %>% min() %>% round_x(., 2)` on this factor, and absolute loadings $\leq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("feeling_distressed", "feeling_lonely", "feeling_frustrated", "feeling_helpless", "feeling_overwhelmed")) %>% gather(factor, loading, -capacity) %>% filter(factor != "MR4") %>% pull(loading) %>% max((abs(.))) %>% round_x(., 2)` on the other three factors. As these numbers suggest, this was a difficult factor to represent, because there were no items that loaded very strongly on the factor (highest loading: `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("feeling_distressed", "feeling_lonely", "feeling_frustrated", "feeling_helpless", "feeling_overwhelmed")) %>% pull("MR4") %>% max() %>% round_x(., 2)`) and because the highest-loading items tended to cross-load on other factors as well. Nonetheless, we decided to try to represent this latent construct because of its theoretical interest and potential clinical importance.

To represent _social connection_, we chose the following items: "feeling excited", "finding something funny", "loving somebody", "learning from other people", and "feeling happy". In the oblimin-transformed `r s1_results$efa$factors`-factor solution from Study 1 all of these items had loadings $\geq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("feeling_excited", "finding_something_funny", "loving_somebody", "learning_from_other_people", "feeling_happy")) %>% pull("MR3") %>% min() %>% round_x(., 2)` on this factor, and absolute loadings $\leq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("feeling_excited", "finding_something_funny", "loving_somebody", "learning_from_other_people", "feeling_happy")) %>% gather(factor, loading, -capacity) %>% filter(factor != "MR3") %>% pull(loading) %>% max(abs(.)) %>% round_x(., 2)` on the other three factors.

Finally, to represent _cognition and control_, we chose the following items: "planning", "having self-control", "reasoning about things", "controlling their emotions", and "telling right from wrong". In the oblimin-transformed `r s1_results$efa$factors`-factor solution from Study 1 all of these items had loadings $\geq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("planning", "having_self_control", "reasoning_about_things", "controlling_their_emotions", "telling_right_from_wrong")) %>% pull("MR1") %>% min() %>% round_x(., 2)` on this factor, and absolute loadings $\leq$ `r s1_results$efa$loadings[] %>% data.frame() %>% rownames_to_column("capacity") %>% filter(capacity %in% c("planning", "having_self_control", "reasoning_about_things", "controlling_their_emotions", "telling_right_from_wrong")) %>% gather(factor, loading, -capacity) %>% filter(factor != "MR1") %>% pull(loading) %>% max(abs(.)) %>% round_x(., 2)` on the other three factors.

Capacities were presented in a random order each time they were displayed (i.e., order varied across both target ages and participants).

#### Target ages and accompanying photographs

The target ages for this study (birth, 4 days, 1 month, 2 months, 4 months, 6 months, 9 months, 12 months, 18 months, 2 years, 3 years, 4 years, and 5 years) were based on the developmental milestone ages described in Study 1. The pairs of photos used to illustrate each target age were identical to those used in Study 1; see Study 1, Methods, for more information. 

As in Study 1, target ages were presented in chronological order, with participants first assessing all 20 capacities for newborns, then 4-day-old infants, then 1-month-old infants, and so forth.

### Analysis plan

#### Preregistered confirmatory analysis of conceptual structure

Based on the results of Study 1, we predicted that the correlational structure of participants' responses would reveal four underlying constructs corresponding to the four factors we identified in Study 1: _bodily sensation_, _negative affect_, _social connection_, and _cognition and control_. To test this hypothesis, we performed an EFA, following the methods established in Study 1. We predicted that, using the same guidelines we applied for conducting and interpreting EFA in Study 1, we would find support for a `r s1_results$efa$factors`-factor solution, with factors that were similar to the factors revealed in that study (e.g., similar groups of capacities that loaded strongly on any given factor).

In line with our predictions, both parallel analysis and Weisman et al.'s factor retention criteria suggested a `r s2_results$efa$factors`-factor solution, which accounted for `r round(100*(s2_results$efa$Vaccounted["Cumulative Var",ncol(s2_results$efa$Vaccounted)]))`% of the total variance in participants' capacity attributions. In this solution, all items loaded most strongly on the factors that they were selected to represent. Minimizing BIC suggested an `r s2_results$efa_bic$factors`-factor solution (accounting for `r round(100*(s2_results$efa_bic$Vaccounted["Cumulative Var",ncol(s2_results$efa_bic$Vaccounted)]))`% of the total variance). 

#### Preregistered confirmatory analysis of perceived developmental trajectories

We preregistered two hypotheses about the perceived developmental trajectories we anticipated to observe in Study 2. (1) We predicted that, on average, participants' perception of children's mental capacities would increase as a function of the target child's age, with participants generally attributing more and greater abilities to older children; and (2) We predicted that, in the domains of _cognition and control_ and _social connection_, participants would perceive relatively more dramatic developmental changes extending later into childhood (i.e., lower intercepts and steeper slopes), whereas in the domains of _bodily sensation_ and _negative affect_ participants would perceive relatively smaller changes across development (i.e., higher intercepts and shallower slopes). 

To test these hypotheses, we conducted a multilevel logistic regression via a generalized additive model, regressing item-level responses onto the domain (i.e., which of the four "factors" this capacity was selected to represented) and a spline-based smooth of the target age of the child fitted separately for each domain, including the maximal random effects structure that would allow the model to converge. Using the "mgcv" package for R [@wood2017], we specified this model as follows: _gam(response ~ s(target age in years, by = domain) + domain, random = list(subject ID = ~1, capacity = ~1)_. Attempting to model random slopes via generalized additive mixed models for domain or for target age resulted in pathological models (e.g., introducing systematic bias in predicted responses). Domain was dummy-coded with _cognition and control_ as the baseline. 

These tests deviated slightly from the preregistered analysis. We originally planned to use "factor scores" (which take into account participants' responses to all 20 items and the relations of those items to all 4 factors) rather than the summary scores described above as our dependent variable for these tests. We planned to derive these factor scores by projecting Study 2 data into the conceptual space defined by the factor solution for Study 1; however, this projection was impossible to implement, because Study 2 included only a subset of the variables used in Study 1. Furthermore, upon reflection, we decided that using any kind of factor scores (even factor scores derived only from Study 2 data) would be inappropriate for testing our hypotheses, which hinge on assessing the _absolute_ degree to which participants attributed different aspects of mental life to children of different ages; factor scores, which function much like z-scores, are instead designed to capture individual differences in responses (i.e., the degree to which a given participant attributed relatively greater or lesser capacities compared to the average of all participants). 

In addition, we note that we originally planned to use a multilevel linear regression to analyze these scores. However, we now consider it misguided to model participants' responses using gaussian distribution, because—both in principle and in our empirical dataset—capacity ratings behaved asymptotically, plateauing near the ceiling of the scale (100, labeled "completely capable") at some point in development. The approach described here—using the binomial distribution—is a far more accurate representation of these responses. Using generalized additive models via the "mgcv" package for R [@wood2017] allowed us to fine-tune the shape of these perceived developmental trajectories via spline-based smoothing (analogous to fitting polynomial terms in the context of linear regression, but without specifying in advance the order of the polynomials included in the model). 

## Study 3

### Participants

`r s3_results$demo %>% distinct() %>% nrow()` US adults participated via MTurk in April 2019. Eligibility requirements were identical to Studies 1-2. Participants were paid $2.90 for approximately 15-20 minutes of their time. As in Study 2, we recruited men and women separately.

According to self-report, participants ranged in age from `r min(s3_results$demo$Age)`-`r max(s3_results$demo$Age)` years ($M$ = `r round_x(mean(s3_results$demo$Age, na.rm = T), 2)` years, $sd$ = `r round_x(sd(s3_results$demo$Age, na.rm = T), 2)` years) and (per our recruitment strategy) included roughly equal numbers of men (`r round((s3_results$demo %>% count(GenderSex) %>% mutate(prop = n/sum(n)) %>% filter(GenderSex == "Male") %>% pull(prop))*100, 0)`%) and women (`r round((s3_results$demo %>% count(GenderSex) %>% mutate(prop = n/sum(n)) %>% filter(GenderSex == "Female") %>% pull(prop))*100, 0)`%). Participants predominantly identified as White (`r round((s3_results$demo %>% count(RaceEthnicity) %>% mutate(prop = n/sum(n)) %>% filter(RaceEthnicity == "White") %>% pull(prop))*100, 0)`%; <`r round(100*(s3_results$demo %>% count(RaceEthnicity) %>% mutate(prop = n/sum(n)) %>% filter(RaceEthnicity != "White") %>% filter(prop == max(prop)) %>% pull(prop)) + 1, 0)`% identified as any other race/ethnicity, identified as more than one race/ethnicity, or declined to disclose). `r round(100*(s3_results$demo %>% count(Education) %>% arrange(desc(Education)) %>% mutate(cumprop = cumsum(n)/sum(n)) %>% filter(grepl("Bachelor", Education)) %>% pull(cumprop)))`% of participants had obtained at least a Bachelor's degree. `r round(100*(s3_results$demo %>% count(Parent) %>% mutate(prop = n/sum(n)) %>% filter(Parent == "Yes") %>% pull(prop)))`% of participants indicated that they were parents.

An additional 79 participants completed the study but were excluded for failing two or more (out of eight) embedded attention checks or for failing to provide a reasonable response in well-formed English sentences to an open-ended comprehension check included at the end of the survey.

### Materials and procedure

After completing a screening questionnaire (in which participants provided information about their gender/sex, age, and English abilities) and providing informed consent, participants were provided with an overview of the study, as follows: "In this study, we will ask you about the mental capacities of children of different ages. We will show you example photos of children at each of the following ages to remind you of what children look like at that age, but please think about children of that age in general when answering these questions. We will ask you about 8 different mental capacities: feeling happy, getting hungry, controlling emotions, feeling distressed, feeling pain, reasoning about things, learning from other people, and feeling helpless. For each mental capacity, we will ask you to assess the extent to which newborns (at birth) have this mental capacity, then 4-day-olds, 1-month-olds, 2-month-olds, 4-month-olds, and so on, ending with 5-year-olds. We'll also ask for your opinion about the different factors that might influence the development of these capacities, such as the child being ‘preprogrammed' to have certain abilities, or the child learning from the people or objects around him or her. At the end of the survey, we will ask you a few demographic questions about yourself. We are interested in your opinions about these questions. Please do not look up extra information about these questions—we just want to know what you think." 

Participants then proceeded through eight trials, in which they assessed each of the eight capacities in turn; the order of capacities was randomized across participants. On each trial, participant assessed children at all 13 target ages used in Study 2 in a fixed order from youngest to oldest. Each age was presented along with a representative pair of photographs (also used in Studies 1 and 2), and participants were asked to answer the following question: "To what extent is a [newborn/4-day-old/1-month-old/etc.] capable of [this capacity]?" Participants responded on a sliding scale from 0 (labeled as "not at all capable") to 100 (labeled as "completely capable").

On each trial, at the bottom of the page participants completed a set of questions about the possible mechanisms that drive development for the capacity in question. First, they were asked, "In your opinion, how important is each of the following factors in the development of children's capacity for [capacity]?" They provided independent ratings on a Likert-type scale from 0 (labeled as "not at all important") to 6 (labeled as "extremely important"), for the following 10 developmental mechanisms: "the child is biologically 'preprogrammed' to have this ability"; "the child has experiences in the womb that give them this ability"; "the child's body grows and matures (for example, muscles get stronger, child gets taller)"; "the child's brain changes (for example, brain grows bigger, more or fewer connections between neurons)"; "the child's senses improve (for example, vision gets sharper, hearing improves)"; "the child observes the objects and the physical world around him or her"; "the child observes the people around him or her"; "the child interacts with the people around him or her"; "people explicitly teach the child how to do this"; "the child actively experiments with how to do this". These items were presented in a fixed order (roughly, from what we perceived to be the most innate or biological mechanisms of development, to what we perceived to be the most learned or social). They were then provided an opportunity to write in "any factors that play a role in the development of children's capacity for [capacity]". Finally, they were asked to choose among these 10 possible mechanisms (or to write in another mechanism) to answer the question, "If you had to choose just one, which of the following factors is the most important in the development of capacities for [capacity]?"
On the final page of the survey, participants were asked to provide additional demographic information.

#### Capacities

Because Study 3 included additional questions probing participants' theories of developmental mechanisms, we reduced the set of mental capacities from 20 in Study 2 to eight in the current study to maintain an appropriate level of participant burden. To select these eight capacities, we drew on the results of Study 2, selecting two items for each of the four factors identified by EFA. For each domain, we selected items that loaded strongly on the relevant factor, did not cross-load strongly on other factors, were sufficiently distinguishable from each other in meaning (i.e., not overly redundant), and captured our qualitative understanding of the domain that each factor corresponded to. We chose "controlling their emotions" and "reasoning about things" to represent the domain of _cognition and control_; "getting hungry" and "feeling pain" to represent the domain of _bodily sensation_; "learning from other people" and "feeling happy" to represent the domain of _social connection_; and "feeling distressed" and "feeling helpless" to represent the domain of _negative affect_. Capacities were presented in a random order for each participant.

#### Target ages and accompanying photographs

The 13 target ages for this study were identical to Study 2, with the exception that we added the parenthetical phrase "(at birth)" when referring to newborns. Each target age was presented with the same pair of photographs used in Study 2. Target ages were presented in chronological order on each trial. 

### Analysis plan

#### Preregistered confirmatory analysis of perceived developmental trajectories

We preregistered two hypotheses about the perceived developmental trajectories we anticipated to observe in Study 3; these were identical to our hypotheses for Study 2 (which were upheld in that study): (1) We predicted that, on average, participants' perception of children's mental capacities would increase as a function of the target child's age, with participants generally attributing more and greater abilities to older children; and (2) We predicted that, in the domains of _cognition and control_ and _social connection_, participants would perceive relatively more dramatic developmental changes extending later into childhood (i.e., lower intercepts and steeper slopes), whereas in the domains of _bodily sensation_ and _negative affect_ participants would perceive relatively smaller changes across development (i.e., higher intercepts and shallower slopes). 

Following Study 2, to test these hypotheses we conducted a multilevel logistic regression via a generalized additive model, regressing item-level responses onto the domain and a spline-based smooth of the target age of the child fitted separately for each domain, including the maximal random effects structure that would allow the model to converge. Using the "mgcv" package for R [@wood2017], we specified this model as follows: _gam(responses ~ s(target age in years, by = domain) + domain + s(subject ID, bs = "re") + s(capacity, bs = "re")_. Again, attempting to model random slopes for domain or for target age resulted in model convergence problems. Again, responses were rescaled to range from 0.005 to 0.995, and domain was dummy-coded with _cognition and control_ as the baseline.

(We note that, in contrast to Study 2, in Study 3 we did preregister this analysis, although we preregistered a linear regression rather than a logistic regression via generalized additive model; see extended notes on our choice of this analysis approach under Study 2.)

#### Preregistered exploratory analysis of perceived developmental trajectories

Beyond the two hypotheses described in the previous section, we also anticipated that perceptions of developmental trajectories in different domains would vary in their curvilinearity—e.g., development in one domain might be perceived to be gradual and linear, while development in another domain might be perceived to occur in stages, at an exponential rate, or to peak at a specific age. To explore these possibilities, we preregistered a regression analysis adding polynomial effects of target age; however, in our reconceptualization of the appropriate way to analyze our data given the distribution of the response variable (see Study 2), this exploratory analysis was superseded by the generalized additive model. 

#### Preregistered exploratory analysis of developmental mechanisms

Finally, we assessed which developmental mechanisms were considered to be the most or least important in the development of each of the individual mental capacities included in this study (*bodily sensation*, *negative affect*, *social connection*, and *cognition and control*). We specified in our preregistration that this assessment would be conducted primarily through visualization, but to complement this visualization we also conducted a multilevel linear regression regressing responses to this question onto developmental mechanism, domain, and an interaction between them, including the maximal random effects structure that allowed the model to converge. Using the "lme4" package for R [@bates.etal2015], we specified this model as follows: _lmer(response ~ mechanism * domain + (1 + mechanism + domain | participant) + (1 | mechanism) + (1 | capacity)_. We also conducted separate, analogous regression analyses for each mechanism separately. In all of these analyses, mechanism and domain were effect-coded for comparisons to the grand mean collapsing across mechanism types and domains.

# Extended results

## Study 1

As described in the main text, we began with an expansive exploration of US adults' attributions of mental life to infants and young children of different ages. We included a wide range of 60 capacities in this initial exploration, from basic physiological sensations, to capacities for perception, cognition, and emotion, to abilities for self-regulation and social interaction. We asked participants to assess these capacities at three different ages (at birth, 9 months, and 5 years), with the goal of characterizing the conceptual structure underlying participants' reasoning about the development of mental life. Our primary analysis for this preregistered study was an exploratory factor analysis (EFA) of participants' capacity ratings. This analysis allowed us to examine which capacities tend to "hang together" in participants' assessments of the mental lives of infants and young children. Following previous work on mind perception, we argue that the suites of capacities revealed by EFA offer a meaningful approximation of the latent conceptual structure underlying participants' reasoning about the developing human mind.

In the main text, we presented the four-factor EFA solution suggested by parallel analysis. Here, we present extended results for this EFA and the alternative EFA solutions suggested by minimizing BIC and by Weisman et al.'s factor retention criteria [@weisman.etal2017a], as well as a visualization of perceptions of development in these mental capacities.

### Exploratory factor analyses (EFAs)

In order to determine how many factors to retain, we examined the results of three factor retention protocols: 

(1) Parallel analysis (presented in the main text), which compares the observed correlation structure to the correlation structure arising from random datasets of the same size; 
(2) Minimizing the Bayesian Information Criterion (BIC), which is one method of optimizing both goodness of fit and parsimony; and 
(3) A set of factor retention criteria that have been used in Weisman et al.'s previous work [@weisman.etal2017a], in which they retained factors with eigenvalues greater than 1.0, which individually accounted for greater than 5% of the shared variance before transformation or rotation, and which were the "dominant" factor (the factor with the strongest absolute factor loading) for at least one mental capacity after transformation.  (In this protocol, a factor had to meet all three of these criteria in order to be retained). 

Our interpretation of how best to characterize our dataset (i.e., how many factors we observe) was guided by the degree of consensus among these three protocols and the interpretability of the retained factors under each protocol. Here we present these three factor retention protocols separately.

All factor analyses were conducted using ordinary least squares to find the minimum residual solution via the "fa" function in the "psych" package for R [@revelle2021].

#### Parallel analysis

```{r s1 parallel analysis}
set.seed(94305)
s1_parallel <- fa.parallel(d1_all, plot = F)
```

```{r s1 efa parallel analysis}
s1_efa_par <- fa(
  d1_all, 
  nfactors = s1_parallel$nfact, 
  rotate = "oblimin")
# s1_efa_par %>% fa.sort()
```

As reported in the main text, parallel analysis suggested retaining `r s1_parallel$nfact` factors. We call these factors _cognition and control_, _social connection_, _bodily sensation_, and _negative affect_. Factor loadings, item complexity, and uniqueness are shown in Supplementary Table \@ref(tab:tabs1efaparload); see also Figure 1 (main text). The amount of total and shared variance accounted for by these factors is shown in Supplementary Table \@ref(tab:tabs1efaparvacc).

It is worth noting that a handful of items loaded relatively strongly on two factors (loadings $\geq$ 0.40), suggesting that these factors are not entirely distinct. The item "feeling distressed", for instance, loaded roughly equally strongly on both _bodily sensation_ and _negative affect_—perhaps reflecting the idea that "distress" might be experienced both as a physical sensation (like pain) and an affective state (like frustration). Likewise, "seeing" loaded equally strongly on both _bodily sensation_ and _social connection_—perhaps reflecting an understanding of vision as both a sensory ability and a critical part of many social interactions (eye contact). The relatively large degree of overlap between _cognition and control_ and _social connection_, in particular, might indicate that participants believed these two constructs to be more strongly connected than other aspects of life; in fact, these two factors were the most strongly correlated of all pairs of factors ($\phi$ = `r round_x(s1_results$efa$Phi["MR1", "MR3"], 2)`). (The next strongest correlation was between _bodily sensation_ and _social connection_: $\phi$  = `r round_x(s1_results$efa$Phi["MR2", "MR3"], 2)`. All other inter-factor correlations: $\phi \le$ `r round_x((s1_results$efa$Phi[upper.tri(s1_results$efa$Phi) == T] %>% sort(decreasing = T))[3], 2)`.)

```{r tabs1efaparload, include = T, warning = F, message = F}
tabs1efaparload <- parameters(s1_efa_par, sort = T) %>% 
  data.frame() %>% 
  rename(Capacity = Variable) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate(across(
    starts_with("MR"), 
    ~ ifelse(. >= 0.6, TRUE, FALSE), 
    .names = "{.col}_bold"
  )) %>%
  rename(`Cog. Ctrl.` = MR1,
         `Soc. Conn.` = MR3,
         `Bod. Sens.` = MR2,
         `Neg. Afft.` = MR4)

tabs1efaparload %>%
  select(-ends_with("_bold")) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "\\text{Study 1, factor solution suggested by parallel analysis: Factor loadings, complexity, and uniqueness.}") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8) %>%
  column_spec(2, width = "1cm", bold = tabs1efaparload[,s1_parallel$nfact+4]) %>%
  column_spec(3, width = "1cm", bold = tabs1efaparload[,s1_parallel$nfact+5]) %>%
  column_spec(4, width = "1cm", bold = tabs1efaparload[,s1_parallel$nfact+6]) %>%
  column_spec(5, width = "1cm", bold = tabs1efaparload[,s1_parallel$nfact+7]) %>%
  footnote(general = "Factor loadings with absolute values >0.60 are in bold.")
```

```{r tabs1efaparvacc, include = T, warning = F, message = F}
s1_efa_par$Vaccounted %>% 
  data.frame() %>% 
  rename(`Cognition and control` = MR1,
         `Social connection` = MR3,
         `Bodily sensation` = MR2,
         `Negative affect` = MR4) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "\\text{Study 1, factor solution suggested by parallel analysis: Variance accounted for.}") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r fig01 code, fig.width = 9, fig.asp = 1.2}
heatmap_fun(s1_efa_par, 
            factor_names = c("Cognition and control",
                             "Bodily sensation",
                             "Social connection",
                             "Negative affect"))

# save plot
if (resave_plots == T | !file.exists("../outputs/fig01.pdf")) {
  ggsave("../outputs/fig01.jpg")
  ggsave("../outputs/fig01.svg")
  ggsave("../outputs/fig01.pdf")
}
```

```{r fig01, fig.cap="Study 1, factor solution suggested by parallel analysis. (Identical to Figure 1 in main text.)"}
# knitr::include_graphics("../outputs/fig01.pdf")
```

#### Minimizing BIC

```{r s1 minimizing bic}
s1_minbic <- VSS(d1_all, plot = F)

s1_minbic <- s1_minbic$vss.stats %>%
  rownames_to_column("nfact") %>%
  mutate(nfact = as.numeric(nfact)) %>%
  filter(BIC == min(BIC)) %>%
  data.frame()
```

```{r s1 efa minimizing bic}
s1_efa_bic <- fa(d1_all, 
                 nfactors = s1_minbic$nfact, 
                 rotate = "oblimin") # does not converge even with max.iter >> 100000; max.iter does affect printout
# s1_efa_bic %>% fa.sort()
```

Minimizing BIC suggested retaining `r s1_minbic$nfact` factors. (Note, however, that this solution does not converge.) Factor loadings, item complexity, and uniqueness are shown in Supplementary Table \@ref(tab:tabs1efabicload). The amount of total and shared variance accounted for by these factors is shown in Supplementary Table \@ref(tab:tabs1efabicvacc).

```{r tabs1efabicload, include = T, warning = F, message = F}
tabs1efabicload <- parameters(s1_efa_bic, sort = T) %>% 
  data.frame() %>% 
  rename(Capacity = Variable) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate(across(
    starts_with("MR"), 
    ~ ifelse(. >= 0.6, TRUE, FALSE), 
    .names = "{.col}_bold"
  ))

tabs1efabicload %>%
  select(-ends_with("_bold")) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "\\text{Study 1, factor solution suggested by minimizing BIC: Factor loadings, complexity, and uniqueness.}") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8) %>%
  column_spec(2, bold = tabs1efabicload[,s1_minbic$nfact+4]) %>%
  column_spec(3, bold = tabs1efabicload[,s1_minbic$nfact+5]) %>%
  column_spec(4, bold = tabs1efabicload[,s1_minbic$nfact+6]) %>%
  column_spec(5, bold = tabs1efabicload[,s1_minbic$nfact+7]) %>%
  column_spec(6, bold = tabs1efabicload[,s1_minbic$nfact+8]) %>%
  column_spec(7, bold = tabs1efabicload[,s1_minbic$nfact+9]) %>%
  footnote(general = "Factor loadings with absolute values >0.60 are in bold.")
```

```{r tabs1efabicvacc, include = T, warning = F, message = F}
s1_efa_bic$Vaccounted %>% 
  data.frame() %>% 
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "\\text{Study 1, factor solution suggested by minimizing BIC: Variance accounted for.}") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

We consider the first of these factors ("`r names(data.frame(s1_efa_bic$loadings[]))[1]`") to resonate with the construct we have called _cognition and control_, the second ("`r names(data.frame(s1_efa_bic$loadings[]))[2]`") to resonate with the construct we have called _bodily sensation_, the third ("`r names(data.frame(s1_efa_bic$loadings[]))[3]`") to resonate with the construct we have called _social connection_, and the fourth ("`r names(data.frame(s1_efa_bic$loadings[]))[4]`") to resonate with the construct we have called _negative affect_. The fifth factor ("`r names(data.frame(s1_efa_bic$loadings[]))[5]`") appears to pick out capacities particularly relevant to social interaction (e.g., social learning, person recognition, humor, listening), and the sixth factor ("`r names(data.frame(s1_efa_bic$loadings[]))[6]`") appears to pick out negative emotions (e.g., anger, annoyance, boredom, social fear); however, there were no capacities that loaded strongly on either of these capacities. 

```{r figS01 code, fig.width = 9, fig.asp = 1.2}
heatmap_fun(s1_efa_bic)

# save plot
if (resave_plots == T | !file.exists("../outputs/figS01.pdf")) {
  ggsave("../outputs/figS01.svg")
  ggsave("../outputs/figS01.pdf")
}
```

```{r figS01, fig.cap="Study 1, factor solution suggested by minimizing BIC.", include = F}
knitr::include_graphics("../outputs/figS01.pdf")
```

#### Weisman et al.'s (2017) factor retention criteria

```{r s1 weisman et al criteria}
s1_weismanetal <- reten_fun(d1_all, rot_type = "oblimin")
s1_weismanetal
```

```{r s1 efa weisman et al criteria}
s1_efa_wdm <- fa(d1_all, 
                 nfactors = s1_weismanetal, 
                 rotate = "oblimin") 
# s1_efa_wdm %>% fa.sort()
```

Weisman et al.'s factor retention criteria suggested retaining `r s1_weismanetal` factors. Factor loadings, item complexity, and uniqueness are shown in Supplementary Table \@ref(tab:tabs1efawdmload). The amount of total and shared variance accounted for by these factors is shown in Supplementary Table \@ref(tab:tabs1efawdmvacc).

```{r tabs1efawdmload, include = T, warning = F, message = F}
tabs1efawdmload <- parameters(s1_efa_wdm, sort = T) %>% 
  data.frame() %>% 
  rename(Capacity = Variable) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate(across(
    starts_with("MR"), 
    ~ ifelse(. >= 0.6, TRUE, FALSE), 
    .names = "{.col}_bold"
  ))

tabs1efawdmload %>%
  select(-ends_with("_bold")) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "\\text{Study 1, factor solution suggested by Weisman et al.'s (2017) factor retention criteria.}") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8) %>%
  column_spec(2, bold = tabs1efawdmload[,s1_weismanetal+4]) %>%
  column_spec(3, bold = tabs1efawdmload[,s1_weismanetal+5]) %>%
  footnote(general = "Factor loadings with absolute values >0.60 are in bold.")
```

```{r tabs1efawdmvacc, include = T, warning = F, message = F}
s1_efa_wdm$Vaccounted %>% 
  data.frame() %>% 
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "Study 1, factor solution suggested by Weisman et al.'s (2017) factor retention criteria: Variance accounted for.") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

We consider the first of these factors ("`r names(data.frame(s1_efa_wdm$loadings[]))[1]`") to resonate most strongly with the construct we have called _cognition and control_, and the second ("`r names(data.frame(s1_efa_wdm$loadings[]))[2]`") to resonate most strongly with the construct we have called _bodily sensation_. These factors are also quite similar to the constructs of "agency" and "experience" in Gray et al.'s original studies of mind perception [@gray.etal2007].

```{r figS02 code, fig.width = 9, fig.asp = 1.2}
heatmap_fun(s1_efa_wdm)

# save plot
if (resave_plots == T | !file.exists("../outputs/figS02.pdf")) {
  ggsave("../outputs/figS02.svg")
  ggsave("../outputs/figS02.pdf")
}
```

```{r figS02, fig.cap="Study 1, factor solution suggested by minimizing BIC."}
knitr::include_graphics("../outputs/figS02.pdf")
```

#### Four-factor solution, varimax rotation

```{r s1 efa varimax rotation}
s1_efa_4_varimax <- fa(d1_all, 
                       nfactors = 4, 
                       rotate = "varimax")
# s1_efa_4_varimax %>% fa.sort()
```

Throughout this manuscript we present EFA results after oblique ("oblimin") transformation. In recent years, oblique transformation has become the preferred approach for many researchers because it allows factors to be correlated with each others (to the extent that these correlations are suggested by the data), thus imposing fewer restrictions on the types of factor structures the might be surfaced by the analysis. 

Here we present one alternative approach, using varimax rotation rather than oblique transformation. This yields a set of orthogonal (uncorrelated) factors.

This rotation is very similar to the oblique transformation that is the focus of our interpretation in the main text. We consider the first of these factors ("`r names(data.frame(s1_efa_4_varimax$loadings[]))[1]`") to resonate with the construct we have called _social connection_, the second ("`r names(data.frame(s1_efa_4_varimax$loadings[]))[2]`") to resonate with the construct we have called _bodily sensation_, the third ("`r names(data.frame(s1_efa_4_varimax$loadings[]))[3]`") to resonate with the construct we have called _cognition and control_, and the fourth ("`r names(data.frame(s1_efa_4_varimax$loadings[]))[4]`") to resonate with the construct we have called _negative affect_. We do note, however, that in this rotation there were no items that loaded most strongly on this fourth factor, in line with the idea that our results provide only moderate evidence for _negative affect_ as a fourth, distinct, domain in participants' representations of the development of mental life.

Factor loadings, item complexity, and uniqueness are shown in Supplementary Table \@ref(tab:tabs1efavarimaxload). The amount of total and shared variance accounted for by these factors is shown in Supplementary Table \@ref(tab:tabs1efavarimaxvacc).

```{r tabs1efavarimaxload, include = T, warning = F, message = F}
tabs1efavarimaxload <- parameters(s1_efa_4_varimax, sort = T) %>% 
  data.frame() %>% 
  rename(Capacity = Variable) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate(across(
    starts_with("MR"), 
    ~ ifelse(. >= 0.6, TRUE, FALSE), 
    .names = "{.col}_bold"
  ))

tabs1efavarimaxload %>%
  select(-ends_with("_bold")) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "\\text{Study 1, four-factor solution after varimax rotation: Factor loadings, complexity, and uniqueness.}") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8) %>%
  column_spec(2, width = "1cm", bold = tabs1efaparload[,s1_parallel$nfact+4]) %>%
  column_spec(3, width = "1cm", bold = tabs1efaparload[,s1_parallel$nfact+5]) %>%
  column_spec(4, width = "1cm", bold = tabs1efaparload[,s1_parallel$nfact+6]) %>%
  column_spec(5, width = "1cm", bold = tabs1efaparload[,s1_parallel$nfact+7]) %>%
  footnote(general = "Factor loadings with absolute values >0.60 are in bold.")
```

```{r tabs1efavarimaxvacc, include = T, warning = F, message = F}
s1_efa_4_varimax$Vaccounted %>% 
  data.frame() %>% 
  rename(`Cognition and control` = MR3,
         `Social connection` = MR1,
         `Bodily sensation` = MR2,
         `Negative affect` = MR4) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "\\text{Study 1, four-factor solution after varimax rotation: Variance accounted for.}") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```


### Capacity ratings: Developmental trajectories

Study 1 also offered a first glimpse into US adults' perceptions of the development of mental life. Supplementary Figure \@ref(fig:figS03) provides a visualization of mean "scores" for each domain of mental life. To calculate these scores, we sorted each of the 60 mental capacities included in this study into on of the four domains (_bodily sensation_, _negative affect_, _social connection_ or _cognition and control_) according to the factor on which it loaded the most strongly and positively (see Supplementary Table \@ref(tab:tabs1efaparload)); for each participant, we then took the average of their responses to capacities in each domain at each target age. As we go on to confirm in Studies 2-3, this visualization suggests that participants perceived capacities for _bodily sensation_ to be substantial at birth and to develop relatively little over the first five years of life; in contrast, they perceived capacities for _negative affect_, _social connection_, and in particular _cognition and control_ to be much more limited at birth and to development dramatically over the first five years of life. 

```{r s1 scales}
d1_scales <- s1_efa_par$loadings[] %>%
  data.frame() %>%
  rownames_to_column("capacity") %>%
  gather(factor, loading, -capacity) %>%
  group_by(capacity) %>%
  top_n(1, loading) %>% #alternatively, could do abs(loading) if there were strong negative loadings
  ungroup() %>%
  mutate(factor = recode_factor(
    factor,
    "MR2" = "Bodily sensation",
    "MR4" = "Negative affect",
    "MR3" = "Social connection",
    "MR1" = "Cognition and control")) %>%
  arrange(factor, desc(loading)) %>%
  mutate(order = 1:nrow(.))
```

```{r s1 cap ratings dataset}
d1_cap_ratings <- d1_all %>%
  rownames_to_column("ResponseId_target") %>%
  separate(ResponseId_target, c("ResponseId", "target")) %>%
  mutate(target_year = recode(
    target,
    target00mo = 0, # newborns
    target09mo = 9/12, # 9 months
    target60mo = 60/12) # 5 years
  ) %>%
  gather(capacity, response, -ResponseId, -starts_with("target")) %>%
  left_join(d1_scales %>% distinct(factor, capacity))
```

```{r s1 cap ratings summary stats}
d1_cap_ratings_sum <- d1_cap_ratings %>%
  group_by(factor, target_year) %>%
  summarise(median = median(response, na.rm = T),
            mean = mean(response, na.rm = T),
            sd = sd(response, na.rm = T)) %>%
  ungroup()

d1_cap_ratings_sum %>%
  arrange(target_year, factor) %>%
  left_join(d1_cap_ratings %>% distinct(target, target_year)) %>%
  select(-target_year) %>%
  select(target, factor, everything()) %>%
  mutate(target = gsub("target", "", target),
         target = gsub("mo", " months", target),
         target = gsub("00 months", "newborn", target),
         target = gsub("^0", "", target)) %>%
  rename(`Target age` = target,
         `Domain` = factor) %>%
  kbl(digits = 2) %>%
  kable_styling() %>%
  collapse_rows(1, valign = "top")
```

```{r s1 gam dataset}
d1_gam <- d1_cap_ratings

contrasts(d1_gam$factor) <- cbind(
  bod_cog = c(1, 0, 0, 0),
  neg_cog = c(0, 1, 0, 0),
  soc_cog = c(0, 0, 1, 0))
```

```{r figS03 code, fig.width = 8, fig.asp = 0.5}
d1_gam %>%
  group_by(ResponseId, target_year, factor) %>%
  # # ALTERNATIVE: take just top 5 items per factor (like Study 2)
  # left_join(d1_scales %>% distinct(factor, capacity, loading)) %>%
  # top_n(5, loading) %>%
  summarise(score = mean(response, na.rm = T)) %>%
  ungroup() %>%
  ggplot(aes(x = target_year, y = score, color = factor)) +
  facet_grid(~ factor) +
  geom_path(aes(group = ResponseId),
            # color = "#160072",
            alpha = 0.2, size = 0.02) +
  geom_path(
    data = . %>%
      group_by(factor, target_year) %>%
      summarise(mean = mean(score, na.rm = T)) %>%
      ungroup,
    aes(y = mean),
    color = "black", lty = 2) +
  geom_pointrange(
    data = . %>%
      group_by(factor, target_year) %>%
      multi_boot_standard(col = "score") %>%
      ungroup,
    aes(y = mean, ymin = ci_lower, ymax = ci_upper),
    color = "black",
    fatten = 1) +
  scale_color_viridis_d(end = 0.9, option = "viridis") +
  scale_y_continuous(breaks = seq(0, 100, 25),
                     labels = seq(0, 100, 25)) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA)) +
  labs(x = "Target age (years)",
       y = "Mean response\n0 = not at all capable, 100 = completely capable") +
  guides(color = guide_none())

# save plot
if (resave_plots == T | !file.exists("../outputs/figS03.pdf")) {
  ggsave("../outputs/figS03.svg")
  ggsave("../outputs/figS03.pdf")
}
```

```{r figS03, fig.cap="Perceived developmental trajectories for four domains of mental life (Study 1). Lighter lines represent individual participants' scores, black points correspond to mean scores across the sample, and error bars are bootstrapped 95% confidence intervals.", include = T}
knitr::include_graphics("../outputs/figS03.pdf")
```

```{r s1 export results}
# save numerical results
s1_results <- list(
  demo = d1_demo,
  efa = s1_efa_par,
  efa_bic = s1_efa_bic,
  efa_wdm = s1_efa_wdm,
  efa_var = s1_efa_4_varimax,
  cap_rat_ss = d1_cap_ratings_sum)

saveRDS(s1_results, "../outputs/s1_results.rds")
```


## Study 2

As described in the main text, in Study 2 we replicated the conceptual structure identified by Study 1 and then used it to chart how different aspects of mental life are perceived to change over development. The design of Study 2 was nearly identical to Study 1, except that instead of assessing 60 capacities for 3 target ages, each participant assessed 20 capacities for 13 target ages.

In the main text, we briefly described the four-factor EFA solution suggested by parallel analysis, and presented the results of a generalized additive model (GAM) predicting participants' responses as a function of target age, domain, and interactions between them. Here, we present the full results for the EFA mentioned in the main text, an alternative EFA solutions suggested by minimizing BIC, and the full results of the GAM.

### Exploratory factor analyses (EFAs)

As in Study 1, we examined the results of three factor retention protocols (parallel analysis, presented in the main text; minimizing BIC; and Weisman et al.'s factor retention criteria [@weisman.etal2017a]), and our interpretation of how best to characterize our dataset was guided by the degree of consensus among these three protocols and the interpretability of the retained factors under each protocol. Here we present these three factor retention protocols separately.

#### Parallel analysis

```{r s2 parallel analysis}
set.seed(94305)
s2_parallel <- fa.parallel(d2_all, plot = F)
```

```{r s2 efa parallel analysis}
s2_efa_par <- fa(d2_all, 
                 nfactors = s2_parallel$nfact, 
                 rotate = "oblimin")
# s2_efa_par %>% fa.sort()
```

As reported in the main text, parallel analysis suggested retaining `r s2_parallel$nfact` factors, and these factors were very similar to those found in Study 1 (_bodily sensation_, _negative affect_, _social connection_, and _cognition and control_); each item loaded most strongly on the factor it was selected to represent. Factor loadings, item complexity, and uniqueness are shown in Supplementary Table \@ref(tab:tabs2efaparload); see also Figure 1 (main text). The amount of total and shared variance accounted for by these factors is shown in Supplementary Table \@ref(tab:tabs2efaparvacc).

```{r tabs2efaparload, include = T, warning = F, message = F}
tabs2efaparload <- parameters(s2_efa_par, sort = T) %>% 
  data.frame() %>% 
  rename(Capacity = Variable) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate(across(
    starts_with("MR"), 
    ~ ifelse(. >= 0.6, TRUE, FALSE), 
    .names = "{.col}_bold"
  )) %>%
  rename(`Neg. Afft.` = MR1,
         `Bod. Sens.` = MR3,
         `Cog. Ctrl.` = MR2,
         `Soc. Conn.` = MR4)

tabs2efaparload %>%
  select(-ends_with("_bold")) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "\\text{Study 2, factor solution suggested by parallel analysis: Factor loadings, complexity, and uniqueness.}") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8) %>%
  column_spec(2, width = "1cm", bold = tabs2efaparload[,s2_parallel$nfact+4]) %>%
  column_spec(3, width = "1cm", bold = tabs2efaparload[,s2_parallel$nfact+5]) %>%
  column_spec(4, width = "1cm", bold = tabs2efaparload[,s2_parallel$nfact+6]) %>%
  column_spec(5, width = "1cm", bold = tabs2efaparload[,s2_parallel$nfact+7]) %>%
  footnote(general = "Factor loadings with absolute values >0.60 are in bold.")
```

```{r tabs2efaparvacc, include = T, warning = F, message = F}
s2_efa_par$Vaccounted %>% 
  data.frame() %>% 
  rename(`Negative affect` = MR1,
         `Bodily sensation` = MR3,
         `Cognition and control` = MR2,
         `Social connection` = MR4) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "\\text{Study 2, factor solution suggested by parallel analysis: Variance accounted for.}") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r figS04 code, fig.width = 9, fig.asp = 0.5}
heatmap_fun(s2_efa_par, 
            factor_names = c("Negative affect",
                             "Cognition and control",
                             "Bodily sensation",
                             "Social connection"))

# save plot
if (resave_plots == T | !file.exists("../outputs/figS04.pdf")) {
  ggsave("../outputs/figS04.svg")
  ggsave("../outputs/figS04.pdf")
}
```

```{r figS04, fig.cap="Study 2, factor solution suggested by parallel analysis. (Identical to Figure 1 in main text.)"}
# knitr::include_graphics("../outputs/figS04.pdf")
```

#### Minimizing BIC

```{r s2 minimizing bic}
s2_minbic <- VSS(d2_all, plot = F)

s2_minbic <- s2_minbic$vss.stats %>%
  rownames_to_column("nfact") %>%
  mutate(nfact = as.numeric(nfact)) %>%
  filter(BIC == min(BIC)) %>%
  data.frame()
```

```{r s2 efa minimizing bic}
s2_efa_bic <- fa(d2_all, 
                 nfactors = s2_minbic$nfact, 
                 rotate = "oblimin") # does not converge even with max.iter >> 100000; max.iter does affect printout
# s2_efa_bic %>% fa.sort()
```

Minimizing BIC suggested retaining `r s2_minbic$nfact` factors. (Note, however, that this solution does not converge.) Factor loadings, item complexity, and uniqueness are shown in Supplementary Table \@ref(tab:tabs2efabicload). The amount of total and shared variance accounted for by these factors is shown in Supplementary Table \@ref(tab:tabs2efabicvacc).

```{r tabs2efabicload, include = T, warning = F, message = F}
tabs2efabicload <- parameters(s2_efa_bic, sort = T) %>% 
  data.frame() %>% 
  rename(Capacity = Variable) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate(Capacity = gsub("_", " ", Capacity)) %>%
  mutate(across(
    starts_with("MR"), 
    ~ ifelse(. >= 0.6, TRUE, FALSE), 
    .names = "{.col}_bold"
  ))

tabs2efabicload %>%
  select(-ends_with("_bold")) %>%
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "\\text{Study 2, factor solution suggested by minimizing BIC: Factor loadings, complexity, and uniqueness.}") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8) %>%
  column_spec(2, bold = tabs2efabicload[,s2_minbic$nfact+4]) %>%
  column_spec(3, bold = tabs2efabicload[,s2_minbic$nfact+5]) %>%
  column_spec(4, bold = tabs2efabicload[,s2_minbic$nfact+6]) %>%
  column_spec(5, bold = tabs2efabicload[,s2_minbic$nfact+7]) %>%
  column_spec(6, bold = tabs2efabicload[,s2_minbic$nfact+8]) %>%
  column_spec(7, bold = tabs2efabicload[,s2_minbic$nfact+9]) %>%
  footnote(general = "Factor loadings with absolute values >0.60 are in bold.")
```

```{r tabs2efabicvacc, include = T, warning = F, message = F}
s2_efa_bic$Vaccounted %>% 
  data.frame() %>% 
  kable("latex", booktabs = T, escape = F, longtable = T, digits = 2,
        caption = "\\text{Study 2, factor solution suggested by minimizing BIC: Variance accounted for.}") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

We consider the first of these factors ("`r names(data.frame(s2_efa_bic$loadings[]))[1]`") to resonate with the construct we have called _cognition and control_, the second ("`r names(data.frame(s2_efa_bic$loadings[]))[2]`") to resonate with the construct we have called _bodily sensation_, the third ("`r names(data.frame(s2_efa_bic$loadings[]))[3]`") to resonate with the construct we have called _social connection_, and the fourth and fifth ("`r names(data.frame(s2_efa_bic$loadings[]))[4]`" and "`r names(data.frame(s2_efa_bic$loadings[]))[5]`") to resonate with the construct we have called _negative affect_. The sixth factor ("`r names(data.frame(s2_efa_bic$loadings[]))[6]`") appears to pick out capacities particularly relevant to social interaction (e.g., social learning, love). We have no strong intuitions about how to interpret the seventh and eighth factors ("`r names(data.frame(s2_efa_bic$loadings[]))[7]`" and "`r names(data.frame(s2_efa_bic$loadings[]))[8]`"). 

```{r figS05 code, fig.width = 9, fig.asp = 1.2}
heatmap_fun(s2_efa_bic)

# save plot
if (resave_plots == T | !file.exists("../outputs/figS05.pdf")) {
  ggsave("../outputs/figS05.svg")
  ggsave("../outputs/figS05.pdf")
}
```

```{r figS05, fig.cap="Study 2, factor solution suggested by minimizing BIC.", include = F}
knitr::include_graphics("../outputs/figS05.pdf")
```

#### Weisman et al.'s (2017) factor retention criteria

```{r s2 weisman et al criteria}
s2_weismanetal <- reten_fun(d2_all, rot_type = "oblimin")
s2_weismanetal
```

```{r s2 efa weisman et al criteria}
s2_efa_wdm <- fa(d2_all, 
                 nfactors = s2_weismanetal, 
                 rotate = "oblimin") 
# s2_efa_wdm %>% fa.sort()
```

Weisman et al.'s factor retention criteria suggested retaining `r s2_weismanetal` factors—i.e., the same `r s2_weismanetal`-factor solution suggested by parallel analysis. (See Supplementary Tables \@ref(tab:tabs2efaparload) and \@ref(tab:tabs2efaparvacc).)

### Capacity ratings: Developmental trajectories

Our primary goal in Study 2 was to chart perceptions of the development of mental life over the first five years. As described in the main text, we preregistered two hypotheses about the perceived developmental trajectories we anticipated to observe in Study 2:

(1) We predicted that, on average, participants' perception of children's mental capacities would increase as a function of the target child's age, with participants generally attributing more and greater abilities to older children.
(2) We predicted that, in the domains of _cognition and control_ and _social connection_, participants would perceive relatively more dramatic developmental changes extending later into childhood (i.e., lower intercepts and steeper slopes), whereas in the domains of _bodily sensation_ and _negative affect_ participants would perceive relatively smaller changes across development (i.e., higher intercepts and shallower slopes). 

To test these hypotheses, we conducted a multilevel logistic regression via a generalized additive model, regressing item-level responses onto the domain (i.e., which of the four "factors" this capacity was selected to represented) and a spline-based smooth of the target age of the child fitted separately for each domain, including the maximal random effects structure that would allow the model to converge. Using the "mgcv" package for R [@wood2017], we specified this model as follows: _gam(response ~ s(target age in years, by = domain) + domain, random = list(subject ID = ~1, capacity = ~1)_. Attempting to model random slopes for domain or for target age resulted in pathological models. Domain was dummy-coded with _cognition and control_ as the baseline. The full results of this model are presented in Supplementary Tables \@ref(tab:tabs2gamfixed) and \@ref(tab:tabs2gamsmooth); see Figure 2, Panel A (main text) for a visualization of these perceived trajectories.

In brief, we view these results as confirming our hypotheses that US adults would rate older children as more capable across the board, but that perceptions of the development of children's capacities would differ across domains. (See main text for extended interpretation.)

```{r s2 scales}
d2_scales <- s2_efa_par$loadings[] %>%
  data.frame() %>%
  rownames_to_column("capacity") %>%
  gather(factor, loading, -capacity) %>%
  group_by(capacity) %>%
  top_n(1, loading) %>%
  ungroup() %>%
  mutate(factor = recode_factor(
    factor,
    "MR3" = "Bodily sensation",
    "MR1" = "Negative affect",
    "MR4" = "Social connection",
    "MR2" = "Cognition and control")) %>%
  arrange(factor, desc(loading)) %>%
  mutate(order = 1:nrow(.))
```

```{r s2 cap ratings dataset}
d2_cap_ratings <- d2_all %>%
  rownames_to_column("ResponseId_target") %>%
  separate(ResponseId_target, c("ResponseId", "target")) %>%
  mutate(target_year = recode(
    target,
    target00mo = 0, # newborns
    target0Xmo = 4/365, # 4 days
    target01mo = 1/12, # 1 month
    target02mo = 2/12, 
    target04mo = 4/12, 
    target06mo = 6/12, 
    target09mo = 9/12,
    target12mo = 12/12,
    target18mo = 18/12,
    target24mo = 24/12,
    target36mo = 36/12,
    target48mo = 48/12,
    target60mo = 60/12)) %>%
  gather(capacity, response, -ResponseId, -starts_with("target")) %>%
  left_join(d2_scales %>% distinct(factor, capacity))
```

```{r s2 cap ratings reliability}
d2_cap_ratings %>%
  unite(ResponseId_target, c(ResponseId, target)) %>%
  distinct(ResponseId_target, factor, capacity, response) %>%
  filter(factor == "Bodily sensation") %>%
  select(-factor) %>%
  spread(capacity, response) %>%
  column_to_rownames("ResponseId_target") %>%
  psych::alpha()

d2_cap_ratings %>%
  unite(ResponseId_target, c(ResponseId, target)) %>%
  distinct(ResponseId_target, factor, capacity, response) %>%
  filter(factor == "Negative affect") %>%
  select(-factor) %>%
  spread(capacity, response) %>%
  column_to_rownames("ResponseId_target") %>%
  psych::alpha()

d2_cap_ratings %>%
  unite(ResponseId_target, c(ResponseId, target)) %>%
  distinct(ResponseId_target, factor, capacity, response) %>%
  filter(factor == "Social connection") %>%
  select(-factor) %>%
  spread(capacity, response) %>%
  column_to_rownames("ResponseId_target") %>%
  psych::alpha()

d2_cap_ratings %>%
  unite(ResponseId_target, c(ResponseId, target)) %>%
  distinct(ResponseId_target, factor, capacity, response) %>%
  filter(factor == "Cognition and control") %>%
  select(-factor) %>%
  spread(capacity, response) %>%
  column_to_rownames("ResponseId_target") %>%
  psych::alpha()
```

```{r s2 cap ratings summary stats}
d2_cap_ratings_sum <- d2_cap_ratings %>%
  group_by(factor, target_year) %>%
  summarise(median = median(response, na.rm = T),
            mean = mean(response, na.rm = T),
            sd = sd(response, na.rm = T)) %>%
  ungroup()

d2_cap_ratings_sum %>%
  arrange(target_year, factor) %>%
  left_join(d2_cap_ratings %>% distinct(target, target_year)) %>%
  select(-target_year) %>%
  select(target, factor, everything()) %>%
  mutate(target = gsub("target", "", target),
         target = gsub("mo", " months", target),
         target = gsub("00 months", "newborn", target),
         target = gsub("0X months", "4 days", target),
         target = gsub("^0", "", target),
         target = gsub("1 months", "1 month", target)) %>%
  rename(`Target age` = target,
         `Domain` = factor) %>%
  kbl(digits = 2) %>%
  kable_styling() %>%
  collapse_rows(1, valign = "top")
```

```{r s2 gam dataset}
d2_gam <- d2_cap_ratings %>%
  # implement rescaling for logistic regression
  mutate(response = response/100) %>%
  mutate(capacity = factor(capacity), 
         ResponseId = factor(ResponseId)) 

contrasts(d2_gam$factor) <- cbind(bod_cog = c(1, 0, 0, 0),
                                  neg_cog = c(0, 1, 0, 0),
                                  soc_cog = c(0, 0, 1, 0))
```

```{r s2 gam}
if (!file.exists("../outputs/s2_gam.RDS")) {
  # takes ~25 minutes to run
  s2_gam <- gam(
    response ~ 
      # fixed effects
      s(target_year, by = factor, k = 12) + factor
    # random intercepts
    + s(capacity, bs = "re") + s(ResponseId, bs = "re")
    ,
    data = d2_gam,
    family = binomial,
    method = "REML",
    niterPQL = 40)
  
  saveRDS(s2_gam, "../outputs/s2_gam.RDS")
  
} else {
  s2_gam <- readRDS("../outputs/s2_gam.RDS")
}

# summary(s2_gam)
# parameters(s2_gam)
# check.gam(s2_gam)
```

```{r tabs2gam}
tabs2gam <- parameters(s2_gam)
```

```{r tabs2gamfixed, include = T}
tabs2gam %>% 
  filter(Component == "conditional") %>%
  select(-c(df, df_error, Component)) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Cognition and control (intercept)",
    grepl("bod", Parameter) ~ "Bodily sensation vs. Cognition and control",
    grepl("neg", Parameter) ~ "Negative affect vs. Cognition and control",
    grepl("soc", Parameter) ~ "Social connection vs. Cognition and control"),
    p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(`Coefficient (log-odds)` = Coefficient) %>%
  kable("latex", booktabs = T, longtable = T, 
        caption = "\\text{Study 2, GAM results: Parametric coefficients (fixed effects).}",
        digits = 2, align = c("lrrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r tabs2gamsmooth, include = T}
tabs2gam %>% 
  filter(Component == "smooth_terms") %>%
  select(-c(Coefficient, SE, starts_with("CI"), df_error, Component)) %>%
  mutate(Parameter = case_when(
    grepl("Bod", Parameter) ~ "Target year: Bodily sensation",
    grepl("Neg", Parameter) ~ "Target year: Negative affect",
    grepl("Soc", Parameter) ~ "Target year: Social connection",
    grepl("Cog", Parameter) ~ "Target year: Cognition and control",
    grepl("capacity", Parameter) ~ "Capacity (random effect)",
    grepl("ResponseId", Parameter) ~ "Participant (random effect)"),
    p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  rename(`Estimated df` = df) %>%
  kable("latex", booktabs = T, longtable = T, 
        caption = "\\text{Study 2, GAM results: Approximate significance of smooth terms.}",
        digits = 2, align = c("lrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r s2 gam predictions}
d2_gam_r1_pred <- d2_gam %>%
  mutate(pred = predict.gam(s2_gam, type = "response", newdata = d2_gam)) %>% 
  group_by(ResponseId, factor, target_year) %>%
  summarise(score = mean(response, na.rm = T),
            pred_score = mean(pred, na.rm = T)) %>%
  ungroup()
```

```{r fig02a_gam, fig.cap="Perceived developmental trajectories for four domains of mental life (Study 2). Lighter lines represent individual participants' scores, black points correspond to mean scores across the sample, error bars are bootstrapped 95% confidence intervals, and thick red lines are predictions from our generalized additive model.", fig.width = 8, fig.asp = 0.5}
d2_gam_r1_rsq <- round_x(summary(s2_gam)$r.sq, 2)

fig02a_gam <- d2_gam_r1_pred %>%
  ggplot(aes(x = target_year, y = score, color = factor)) +
  facet_grid(~ factor) +
  geom_path(aes(group = ResponseId),
            alpha = 0.2, size = 0.02) +
  geom_smooth(aes(y = pred_score), 
              color = "#f94802",
              fill = NA, size = 1, lty = 1) +
  geom_pointrange(
    data = . %>%
      group_by(factor, target_year) %>%
      multi_boot_standard(col = "score") %>%
      ungroup,
    aes(y = mean, ymin = ci_lower, ymax = ci_upper),
    color = "black",
    fatten = 1) +
  scale_color_viridis_d(end = 0.9, option = "viridis") +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = seq(0, 100, 25)) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA)) +
  labs(x = "Target age (years)",
       y = "Mean response", # \n0 = not at all capable, 100 = completely capable",
       caption = bquote(paste("Adjusted R"^"2"*" = ", .(d2_gam_r1_rsq)))) +
  guides(color = guide_none())

# make jpg version (darker)
fig02a_gam_jpg <- d2_gam_r1_pred %>%
  ggplot(aes(x = target_year, y = score, color = factor)) +
  facet_grid(~ factor) +
  geom_path(aes(group = ResponseId),
            alpha = 1, size = 0.02) +
  geom_smooth(aes(y = pred_score), 
              color = "#f94802",
              fill = NA, size = 1, lty = 1) +
  geom_pointrange(
    data = . %>%
      group_by(factor, target_year) %>%
      multi_boot_standard(col = "score") %>%
      ungroup,
    aes(y = mean, ymin = ci_lower, ymax = ci_upper),
    color = "black",
    fatten = 1) +
  scale_color_viridis_d(end = 0.9, option = "viridis") +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = seq(0, 100, 25)) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA)) +
  labs(x = "Target age (years)",
       y = "Mean response", # \n0 = not at all capable, 100 = completely capable",
       caption = bquote(paste("Adjusted R"^"2"*" = ", .(d2_gam_r1_rsq)))) +
  guides(color = guide_none())

# # save plot
# if (resave_plots == T | !file.exists("../outputs/fig02a_gam.pdf")) {
#   ggsave("../outputs/fig02a_gam.svg")
#   ggsave("../outputs/fig02a_gam.pdf")
#   ggsave("../outputs/fig02a_gam.jpg")
# }
```

```{r s2 export results}
# save numerical results
s2_results <- list(
  demo = d2_demo,
  efa = s2_efa_par,
  efa_bic = s2_efa_bic,
  efa_wdm = s2_efa_wdm,
  cap_rat_ss = d2_cap_ratings_sum,
  gam = s2_gam)

saveRDS(s2_results, "../outputs/s2_results.rds")
```


## Study 3

As described in the main text, Study 3 was designed with two goals in mind: First, we aimed to provide a more direct assessment of the differences in perceived developmental trajectories surfaced by Study 2. Second, we aimed to probe the intuitive theories that might underlie the differentiation of the four factors surfaced by Studies 1-2 and the perceived developmental trajectories surfaced by Study 2 (and the current study).

In the main text, we presented the results of a generalized additive model (GAM) predicting participants' capacity ratings as a function of target age, domain, and interactions between them; and briefly presented the results of a series of multilevel linear regression models predicting participants' ratings of developmental mechanisms as a function of domain; and provided visualizations of participants' selections of the "most important" mechanism for each capacity, by domain. Here, we present the full results of all of these models, as well as the results of a multilevel logistic regression model of the selection of "most important" mechanism.

### Capacity ratings: Developmental trajectories

We preregistered two hypotheses about the perceived developmental trajectories we anticipated to observe in Study 3 (identical to our hypotheses for Study 2):

(1) We predicted that, on average, participants' perception of children's mental capacities would increase as a function of the target child's age, with participants generally attributing more and greater abilities to older children.
(2) We predicted that, in the domains of _cognition and control_ and _social connection_, participants would perceive relatively more dramatic developmental changes extending later into childhood (i.e., lower intercepts and steeper slopes), whereas in the domains of _bodily sensation_ and _negative affect_ participants would perceive relatively smaller changes across development (i.e., higher intercepts and shallower slopes). 

Following Study 2, to test these hypotheses we conducted a multilevel logistic regression via a generalized additive model, regressing item-level responses onto the domain and a spline-based smooth of the target age of the child fitted separately for each domain, including the maximal random effects structure that would allow the model to converge. Using the "mgcv" package for R [@wood2017], we specified this model as follows: _gam(responses ~ s(target age in years, by = domain) + domain + s(subject ID, bs = "re") + s(capacity, bs = "re")_. Again, attempting to model random slopes for domain or for target age resulted in pathological models. Again, domain was dummy-coded with _cognition and control_ as the baseline. The full results of this model are presented in Supplementary Tables \@ref(tab:tabs3gamfixed) and \@ref(tab:tabs3gamsmooth); see Figure 2, Panel B (main text) for a visualization of these perceived trajectories.

In brief, we view these results as confirming our hypotheses that US adults would rate older children as more capable across the board, but that perceptions of the development of children's capacities would differ across domains. (See main text for extended interpretation.)

```{r s3 cap ratings dataset}
d3_cap_ratings <- d3 %>%
  mutate(target_year = target_num/12) %>%
  mutate(factor = recode_factor(
    domain,
    "BOD" = "Bodily sensation",
    "NEG" = "Negative affect",
    "POS" = "Social connection",
    "COG" = "Cognition and control")) %>%
  select(-domain)
```

```{r s3 cap ratings reliability}
d3_cap_ratings %>%
  unite(ResponseId_target, c(ResponseId, target)) %>%
  distinct(ResponseId_target, factor, capacity, response) %>%
  filter(factor == "Bodily sensation") %>%
  select(-factor) %>%
  spread(capacity, response) %>%
  column_to_rownames("ResponseId_target") %>%
  psych::alpha()

d3_cap_ratings %>%
  unite(ResponseId_target, c(ResponseId, target)) %>%
  distinct(ResponseId_target, factor, capacity, response) %>%
  filter(factor == "Negative affect") %>%
  select(-factor) %>%
  spread(capacity, response) %>%
  column_to_rownames("ResponseId_target") %>%
  psych::alpha()

d3_cap_ratings %>%
  unite(ResponseId_target, c(ResponseId, target)) %>%
  distinct(ResponseId_target, factor, capacity, response) %>%
  filter(factor == "Social connection") %>%
  select(-factor) %>%
  spread(capacity, response) %>%
  column_to_rownames("ResponseId_target") %>%
  psych::alpha()

d3_cap_ratings %>%
  unite(ResponseId_target, c(ResponseId, target)) %>%
  distinct(ResponseId_target, factor, capacity, response) %>%
  filter(factor == "Cognition and control") %>%
  select(-factor) %>%
  spread(capacity, response) %>%
  column_to_rownames("ResponseId_target") %>%
  psych::alpha()
```

```{r s3 cap ratings summary stats}
d3_cap_ratings_sum <- d3_cap_ratings %>%
  group_by(factor, target_year) %>%
  summarise(median = median(response, na.rm = T),
            mean = mean(response, na.rm = T),
            sd = sd(response, na.rm = T)) %>%
  ungroup() %>%
  arrange(target_year, factor)

d3_cap_ratings_sum %>%
  arrange(target_year, factor) %>%
  left_join(d3_cap_ratings %>% distinct(target, target_year)) %>%
  select(-target_year) %>%
  select(target, factor, everything()) %>%
  mutate(target = gsub("target", "", target),
         target = gsub("mo", " months", target),
         target = gsub("00 months", "newborn", target),
         target = gsub("0X months", "4 days", target),
         target = gsub("^0", "", target),
         target = gsub("1 months", "1 month", target)) %>%
  rename(`Target age` = target,
         `Domain` = factor) %>%
  kbl(digits = 2) %>%
  kable_styling() %>%
  collapse_rows(1, valign = "top")
```

```{r s3 gam dataset}
d3_gam <- d3_cap_ratings %>%
  # implement rescaling for logistic regression
  mutate(response = response/100) %>%
  mutate(capacity = factor(capacity),
         ResponseId = factor(ResponseId)) 

contrasts(d3_gam$factor) <- cbind(bod_cog = c(1, 0, 0, 0),
                                  neg_cog = c(0, 1, 0, 0),
                                  soc_cog = c(0, 0, 1, 0))
```

```{r s3 gam}
if (!file.exists("../outputs/s3_gam.RDS")) {
  # takes ~16 minutes to run
  s3_gam <- gam(
    response ~ 
      # fixed effects
      s(target_year, by = factor, k = 12) + factor
    # random intercepts
    + s(capacity, bs = "re") + s(ResponseId, bs = "re")
    ,
    data = d3_gam,
    family = binomial,
    method = "REML",
    niterPQL = 40)
  
  saveRDS(s3_gam, "../outputs/s3_gam.RDS")
  
} else {
  s3_gam <- readRDS("../outputs/s3_gam.RDS")
}

# summary(s3_gam)
```

```{r tabs3gam}
tabs3gam <- parameters(s3_gam)
```

```{r tabs3gamfixed, include = T}
tabs3gam %>% 
  filter(Component == "conditional") %>%
  select(-c(df, df_error, Component)) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Cognition and control (intercept)",
    grepl("bod", Parameter) ~ "Bodily sensation vs. Cognition and control",
    grepl("neg", Parameter) ~ "Negative affect vs. Cognition and control",
    grepl("soc", Parameter) ~ "Social connection vs. Cognition and control"),
    p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(`Coefficient (log-odds)` = Coefficient) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "\\text{Study 3, GAM results: Parametric coefficients (fixed effects).}",
        digits = 2, align = c("lrrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r tabs3gamsmooth, include = T}
tabs3gam %>% 
  filter(Component == "smooth_terms") %>%
  select(-c(Coefficient, SE, starts_with("CI"), df_error, Component)) %>%
  mutate(
    Parameter = case_when(
      grepl("Bod", Parameter) ~ "Target year: Bodily sensation",
      grepl("Neg", Parameter) ~ "Target year: Negative affect",
      grepl("Soc", Parameter) ~ "Target year: Social connection",
      grepl("Cog", Parameter) ~ "Target year: Cognition and control",
      grepl("capacity", Parameter) ~ "Capacity (random effect)",
      grepl("ResponseId", Parameter) ~ "Participant (random effect)"),
    p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  rename(`Estimated df` = df) %>%
  kable("latex", booktabs = T, longtable = T, 
        caption = "\\text{Study 3, GAM results: Approximate significance of smooth terms.}",
        digits = 2, align = c("lrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

```{r s3 gam predictions}
d3_gam_r1_pred <- d3_gam %>%
  mutate(pred = predict.gam(s3_gam, type = "response", newdata = d3_gam)) %>% 
  group_by(ResponseId, factor, target_year) %>%
  summarise(score = mean(response, na.rm = T),
            pred_score = mean(pred, na.rm = T)) %>%
  ungroup()
```

```{r fig02b_gam, fig.cap="Perceived developmental trajectories for four domains of mental life (Study 3). Scores were calculated by averaging a participant's responses to the two capacity items for each domain of mental life. Lighter lines represent individual participants' scores, black points correspond to mean scores across the sample, error bars are bootstrapped 95% confidence intervals, and thick red lines are predictions from our generalized additive model.", fig.width = 8, fig.asp = 0.5}
d3_gam_r1_rsq <- round_x(summary(s3_gam)$r.sq, 2)

fig02b_gam <- d3_gam_r1_pred %>%
  ggplot(aes(x = target_year, y = score, color = factor)) +
  facet_grid(~ factor) +
  geom_path(aes(group = ResponseId),
            alpha = 0.2, size = 0.02) +
  geom_smooth(aes(y = pred_score), 
              color = "#f94802",
              fill = NA, size = 1, lty = 1) +
  geom_pointrange(
    data = . %>%
      group_by(factor, target_year) %>%
      multi_boot_standard(col = "score") %>%
      ungroup,
    aes(y = mean, ymin = ci_lower, ymax = ci_upper),
    color = "black",
    fatten = 1) +
  scale_color_viridis_d(end = 0.9, option = "viridis") +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = seq(0, 100, 25)) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA)) +
  labs(x = "Target age (years)",
       y = "Mean response", # \n0 = not at all capable, 100 = completely capable",
       caption = bquote(paste("Adjusted R"^"2"*" = ", .(d3_gam_r1_rsq)))) +
  guides(color = guide_none())

# make jpg version (darker)
fig02b_gam_jpg <- d3_gam_r1_pred %>%
  ggplot(aes(x = target_year, y = score, color = factor)) +
  facet_grid(~ factor) +
  geom_path(aes(group = ResponseId),
            alpha = 1, size = 0.02) +
  geom_smooth(aes(y = pred_score), 
              color = "#f94802",
              fill = NA, size = 1, lty = 1) +
  geom_pointrange(
    data = . %>%
      group_by(factor, target_year) %>%
      multi_boot_standard(col = "score") %>%
      ungroup,
    aes(y = mean, ymin = ci_lower, ymax = ci_upper),
    color = "black",
    fatten = 1) +
  scale_color_viridis_d(end = 0.9, option = "viridis") +
  scale_y_continuous(breaks = seq(0, 1, 0.25),
                     labels = seq(0, 100, 25)) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA)) +
  labs(x = "Target age (years)",
       y = "Mean response", # \n0 = not at all capable, 100 = completely capable",
       caption = bquote(paste("Adjusted R"^"2"*" = ", .(d3_gam_r1_rsq)))) +
  guides(color = guide_none())

# # save plot
# if (resave_plots == T | !file.exists("../outputs/fig02b_gam.pdf")) {
#   ggsave("../outputs/fig02b_gam.svg")
#   ggsave("../outputs/fig02b_gam.pdf")
# }
```

#### Figure 2

```{r fig02, fig.cap="Perceived developmental trajectories for four domains of mental life (Studies 2-3). Lighter lines represent individual participants' responses, black points correspond to mean responses across the sample, error bars are bootstrapped 95% confidence intervals, and thick red lines are predictions from our generalized additive models. In Study 2 (Panel A), participants assessed 5 capacities within each domain, and assessed all capacities for a given target age before moving on to the next target age. In Study 3 (Panel B), participants assessed 2 capacities within each domain, and assessed a single capacity for all target ages before moving on to the next capacity.", fig.width = 8, fig.asp = 0.7}
plot_grid(fig02a_gam, fig02b_gam,
          ncol = 1, labels = "AUTO")

# save plot
if (resave_plots == T | !file.exists("../outputs/fig02.pdf")) {
  ggsave("../outputs/fig02.svg")
  ggsave("../outputs/fig02.pdf")
  ggsave("../outputs/fig02.jpg")
}
```

```{r fig02_jpg, fig.cap="Perceived developmental trajectories for four domains of mental life (Studies 2-3). Lighter lines represent individual participants' responses, black points correspond to mean responses across the sample, error bars are bootstrapped 95% confidence intervals, and thick red lines are predictions from our generalized additive models. In Study 2 (Panel A), participants assessed 5 capacities within each domain, and assessed all capacities for a given target age before moving on to the next target age. In Study 3 (Panel B), participants assessed 2 capacities within each domain, and assessed a single capacity for all target ages before moving on to the next capacity.", fig.width = 8, fig.asp = 0.7}
plot_grid(fig02a_gam_jpg, fig02b_gam_jpg,
          ncol = 1, labels = "AUTO")

# save plot   
if (resave_plots == T | !file.exists("../outputs/fig02.jpg")) {
  ggsave("../outputs/fig02.jpg")
}
```

### Developmental mechanisms: Ratings

```{r s3 mech ratings datasets a}
d3_mech_ratings_wide <- d3_dev_factor_rating %>%
  select(-domain) %>%
  spread(dev_factor, response) %>%
  unite(RepsonseId_capacity, c(ResponseId, capacity)) %>%
  column_to_rownames("RepsonseId_capacity")
```

```{r s3 mech ratings datasets b}
d3_mech_ratings <- d3_dev_factor_rating %>%
  mutate(factor = recode_factor(
    domain,
    "BOD"= "Bodily sensation",
    "NEG" = "Negative affect",
    "POS" = "Social connection",
    "COG" = "Cognition and control")) %>%
  select(-domain) %>%
  mutate(
    dev_factor_text = recode_factor(
      dev_factor,
      "preprogrammed" = "biologically 'preprogrammed'",
      "womb_experiences" = "experiences in the womb",
      "body_grows" = "body grows and matures",
      "brain_changes" = "brain changes",
      "senses_improve" = "senses improve",
      "observes_objects" = "observes objects and physical world",
      "experiments" = "actively experiments",
      "observes_people" = "observes people",
      "interacts_people" = "interacts with people",
      "people_teach" = "people explicitly teach"))

contrasts(d3_mech_ratings$factor) <- cbind(bod_gm = c(1, -1, 0, 0),
                                           soc_gm = c(0, -1, 1, 0),
                                           cog_gm = c(0, -1, 0, 1))

# contrasts(d3_mech_ratings$factor) <- cbind(bod_gm = c(1, 0, -1, 0),
#                                            neg_gm = c(0, 1, -1, 0),
#                                            cog_gm = c(0, 0, -1, 1))

contrasts(d3_mech_ratings$dev_factor) <- cbind(
  # baseline/drop: senses improve
  prep_gm = c(1, 0, 0, 0, -1, 0, 0, 0, 0, 0),
  womb_gm = c(0, 1, 0, 0, -1, 0, 0, 0, 0, 0),
  body_gm = c(0, 0, 1, 0, -1, 0, 0, 0, 0, 0),
  brai_gm = c(0, 0, 0, 1, -1, 0, 0, 0, 0, 0),
  obso_gm = c(0, 0, 0, 0, -1, 1, 0, 0, 0, 0),
  obsp_gm = c(0, 0, 0, 0, -1, 0, 1, 0, 0, 0),
  intp_gm = c(0, 0, 0, 0, -1, 0, 0, 1, 0, 0),
  teac_gm = c(0, 0, 0, 0, -1, 0, 0, 0, 1, 0),
  expt_gm = c(0, 0, 0, 0, -1, 0, 0, 0, 0, 1))
```

#### Summary tables

```{r s3 mech ratings summary stats}
# by mechanism (no domain)
d3_mech_ratings_sum_a <- d3_mech_ratings %>%
  # participant-level mean for each mechanism
  group_by(ResponseId, dev_factor_text) %>%
  summarise(score_a = mean(response, na.rm = T)) %>%
  ungroup() %>%
  # group-level mean of means for each mechanism 
  group_by(dev_factor_text) %>%
  summarise(median = median(score_a, na.rm = T),
            mean = mean(score_a, na.rm = T),
            sd = sd(score_a, na.rm = T)) %>%
  ungroup() %>%
  arrange(dev_factor_text) %>%
  rename(`Developmental mechanism` = dev_factor_text)

d3_mech_ratings_sum_a %>%
  kbl(digits = 2) %>% 
  kable_styling() %>%
  collapse_rows(1, valign = "top")

# by domain (no mechanism)
d3_mech_ratings_sum_b <- d3_mech_ratings %>%
  # participant-level mean for each domain
  group_by(ResponseId, factor) %>%
  summarise(score_a = mean(response, na.rm = T)) %>%
  ungroup() %>%
  # group-level mean of means for each domain 
  group_by(factor) %>%
  summarise(median = median(score_a, na.rm = T),
            mean = mean(score_a, na.rm = T),
            sd = sd(score_a, na.rm = T)) %>%
  ungroup() %>%
  arrange(factor) %>%
  rename(Domain = factor)

d3_mech_ratings_sum_b %>%
  kbl(digits = 2) %>% 
  kable_styling() %>%
  collapse_rows(1, valign = "top")

# by mechanism and domain
d3_mech_ratings_sum_c <- d3_mech_ratings %>%
  # participant-level mean for each mechanism, for each domain
  group_by(ResponseId, factor, dev_factor_text) %>%
  summarise(score_a = mean(response, na.rm = T)) %>%
  ungroup() %>%
  # group-level mean of means for each domain 
  group_by(factor, dev_factor_text) %>%
  summarise(median = median(score_a, na.rm = T),
            mean = mean(score_a, na.rm = T),
            sd = sd(score_a, na.rm = T)) %>%
  ungroup() %>%
  arrange(factor, dev_factor_text) %>%
  rename(Domain = factor,
         `Developmental mechanism` = dev_factor_text)

d3_mech_ratings_sum_c %>%
  kbl(digits = 2) %>% 
  kable_styling() %>%
  collapse_rows(1, valign = "top")

# by mechanism, domain, and capacity
d3_mech_ratings_sum_d <- d3_mech_ratings %>%
  # participant-level mean for each mechanism, for each domain
  group_by(ResponseId, factor, capacity, dev_factor_text) %>%
  summarise(score_a = mean(response, na.rm = T)) %>%
  ungroup() %>%
  # group-level mean of means for each domain 
  group_by(factor, capacity, dev_factor_text) %>%
  summarise(median = median(score_a, na.rm = T),
            mean = mean(score_a, na.rm = T),
            sd = sd(score_a, na.rm = T)) %>%
  ungroup() %>%
  arrange(factor, capacity, dev_factor_text) %>%
  mutate(capacity = gsub("_", " ", capacity)) %>%
  rename(Domain = factor,
         Capacity = capacity,
         `Developmental mechanism` = dev_factor_text)

d3_mech_ratings_sum_d %>%
  kbl(digits = 2) %>% 
  kable_styling() %>%
  collapse_rows(1:2, valign = "top")
```

#### Regression models

To supplement our visual assessment of which developmental mechanisms were considered to be the most or least important in the development of each of the individual mental capacities included in this study (*bodily sensation*, *negative affect*, *social connection*, and *cognition and control*), we conducted a series of multilevel linear regression models, regressing responses to this question onto domain, including the maximal random effects structure that allowed the model to converge. Using the "lme4" package for R [@bates.etal2015], we specified these models as follows: _lmer(response ~ domain + (1 | participant) + (1 | capacity)_. In these analyses, domain and developmental mechanism were both effect-coded for comparisons to the grand mean collapsing across domains and mechanisms.

The full results of these models are presented in Supplementary Tables \@ref(tab:tabs3lmerall), \@ref(tab:tabs3lmerbod), \@ref(tab:tabs3lmerneg), \@ref(tab:tabs3lmersoc), and \@ref(tab:tabs3lmercog); see also Figure 3 (main text).

##### All domains

In brief, this analysis suggested that the perceived importance of these developmental mechanisms varied both overall (see main effects) and across domains (see interaction terms).

```{r s3 mech ratings lmer}
s3_lmer <- lmer(response ~ dev_factor * factor + 
                  (1 | ResponseId) + 
                  (1 | capacity),
                control = lmerControl(optimizer = "bobyqa"),
                data = d3_mech_ratings)

# summary(s3_lmer)
# parameters(s3_lmer, standardize = "refit")
```

```{r tabs3lmerall, results = "asis", include = T, message = F, warning = F}
cat("\\begin{landscape}")
parameters(s3_lmer, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, t, df_error, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("dev_factorprep_gm", "Mechanism (Biologically 'preprogrammed' vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorwomb_gm", "Mechanism (Experiences in the womb vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorbody_gm", "Mechanism (Body grows and matures vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorbrai_gm", "Mechanism (Brain changes vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorsens_gm", "Mechanism (Senses improve vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorobso_gm", "Mechanism (Observes objects and physical world vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorexpt_gm", "Mechanism (Actively experiments vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorobsp_gm", "Mechanism (Observes people vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorintp_gm", "Mechanism (Interacts with people vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorteac_gm", "Mechanism (People explicitly teach vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensation (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(Parameter = str_wrap(Parameter, 80)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(df = df_error) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "Study 3, multilevel linear model results (all domains): Fixed effects.",
        digits = 2, align = c("lrrrrrrrrrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 7.5)
cat("\\end{landscape}")
```

##### Bodily sensation

For capacities in the domain of _bodily sensation_, this analysis suggested that participants rated biological preprogramming, body changes, and brain changes as more important drivers of development than other mechanisms; see also Figure 3 (main text).

```{r s3 mech ratings lmer bod}
s3_lmer_bod <- lmer(response ~ dev_factor + 
                      (1 | ResponseId) + 
                      (1 | capacity),
                    control = lmerControl(optimizer = "bobyqa"),
                    data = d3_mech_ratings %>% 
                      filter(factor == "Bodily sensation"))
# summary(s3_lmer_bod)
# parameters(s3_lmer_bod, standardize = "refit")
```

```{r tabs3lmerbod, include = T, message = F, warning = F}
parameters(s3_lmer_bod, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, t, df_error, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("dev_factorprep_gm", "Mechanism (Biologically 'preprogrammed' vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorwomb_gm", "Mechanism (Experiences in the womb vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorbody_gm", "Mechanism (Body grows and matures vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorbrai_gm", "Mechanism (Brain changes vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorsens_gm", "Mechanism (Senses improve vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorobso_gm", "Mechanism (Observes objects and physical world vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorexpt_gm", "Mechanism (Actively experiments vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorobsp_gm", "Mechanism (Observes people vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorintp_gm", "Mechanism (Interacts with people vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorteac_gm", "Mechanism (People explicitly teach vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensation (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(df = df_error) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "\\text{Study 3, multilevel linear model results (bodily sensation): Fixed effects.}",
        digits = 2, align = c("lrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

##### Negative affect

For capacities in the domain of _negative affect_, this analysis suggested that participants rated biological preprogramming, brain changes, social interactions, and observations of other people as more important drivers of development than other mechanisms; see also Figure 3 (main text).

```{r s3 mech ratings lmer neg}
s3_lmer_neg <- lmer(response ~ dev_factor + 
                      (1 | ResponseId) + 
                      (1 | capacity),
                    control = lmerControl(optimizer = "bobyqa"),
                    data = d3_mech_ratings %>% 
                      filter(factor == "Negative affect"))

# summary(s3_lmer_neg)
# parameters(s3_lmer_neg, standardize = "refit")
```

```{r tabs3lmerneg, include = T, message = F, warning = F}
parameters(s3_lmer_neg, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, t, df_error, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("dev_factorprep_gm", "Mechanism (Biologically 'preprogrammed' vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorwomb_gm", "Mechanism (Experiences in the womb vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorbody_gm", "Mechanism (Body grows and matures vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorbrai_gm", "Mechanism (Brain changes vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorsens_gm", "Mechanism (Senses improve vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorobso_gm", "Mechanism (Observes objects and physical world vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorexpt_gm", "Mechanism (Actively experiments vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorobsp_gm", "Mechanism (Observes people vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorintp_gm", "Mechanism (Interacts with people vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorteac_gm", "Mechanism (People explicitly teach vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensation (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(df = df_error) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "\\text{Study 3, multilevel linear model results (negative affect): Fixed effects.}",
        digits = 2, align = c("lrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

##### Social connection

For capacities in the domain of _social connection_, this analysis suggested that participants rated social interactions, observations of other people, and brain changes as more important drivers of development than other mechanisms; see also Figure 3 (main text).

```{r s3 mech ratings lmer soc}
s3_lmer_soc <- lmer(response ~ dev_factor + 
                      (1 | ResponseId) + 
                      (1 | capacity),
                    control = lmerControl(optimizer = "bobyqa"),
                    data = d3_mech_ratings %>% 
                      filter(factor == "Social connection"))

# summary(s3_lmer_soc)
# parameters(s3_lmer_soc)
```

```{r tabs3lmersoc, include = T, message = F, warning = F}
parameters(s3_lmer_soc, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, t, df_error, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("dev_factorprep_gm", "Mechanism (Biologically 'preprogrammed' vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorwomb_gm", "Mechanism (Experiences in the womb vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorbody_gm", "Mechanism (Body grows and matures vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorbrai_gm", "Mechanism (Brain changes vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorsens_gm", "Mechanism (Senses improve vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorobso_gm", "Mechanism (Observes objects and physical world vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorexpt_gm", "Mechanism (Actively experiments vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorobsp_gm", "Mechanism (Observes people vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorintp_gm", "Mechanism (Interacts with people vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorteac_gm", "Mechanism (People explicitly teach vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensation (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(df = df_error) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "\\text{Study 3, multilevel linear model results (social connection): Fixed effects.}",
        digits = 2, align = c("lrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```

##### Cognition and control

For capacities in the domain of _cognition and control_, this analysis suggested that participants rated brain changes, social interactions, observations of other people, explicit teaching, and active experimentation as more important drivers of development than other mechanisms; see also Figure 3 (main text).

```{r s3 mech ratings lmer cog}
s3_lmer_cog <- lmer(response ~ dev_factor + 
                      (1 | ResponseId) + 
                      (1 | capacity),
                    control = lmerControl(optimizer = "bobyqa"),
                    data = d3_mech_ratings %>% 
                      filter(factor == "Cognition and control"))

# summary(s3_lmer_cog)
# parameters(s3_lmer, standardize = "refit")
```

```{r tabs3lmercog, include = T, message = F, warning = F}
parameters(s3_lmer_cog, standardize = "refit") %>%
  filter(Effects == "fixed") %>%
  select(Parameter, Coefficient, SE, CI_low, CI_high, t, df_error, p) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ Parameter)) %>%
  mutate(Parameter = gsub("dev_factorprep_gm", "Mechanism (Biologically 'preprogrammed' vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorwomb_gm", "Mechanism (Experiences in the womb vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorbody_gm", "Mechanism (Body grows and matures vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorbrai_gm", "Mechanism (Brain changes vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorsens_gm", "Mechanism (Senses improve vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorobso_gm", "Mechanism (Observes objects and physical world vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorexpt_gm", "Mechanism (Actively experiments vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorobsp_gm", "Mechanism (Observes people vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorintp_gm", "Mechanism (Interacts with people vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("dev_factorteac_gm", "Mechanism (People explicitly teach vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorbod_gm", "Bodily sensation (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorsoc_gm", "Social connection (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("factorcog_gm", "Cognition and control (vs. GM)", Parameter)) %>%
  mutate(Parameter = gsub("\\:", " x ", Parameter)) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Parameter, Coefficient, `95% CI`, everything()) %>%
  rename(df = df_error) %>%
  kable("latex", booktabs = T, longtable = T,
        caption = "\\text{Study 3, multilevel linear model results (cognition and control): Fixed effects.}",
        digits = 2, align = c("lrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8)
```


### Developmental mechanisms: Selection of "most important"

As described in the main text, we also assessed participants' responses to forced-choice questions about the "most important" factor driving the development of each of these capacities. 

```{r s3 mech choice datasets}
d3_mech_choice <- d3_dev_factor_most_important_choice %>%
  mutate(
    factor = recode_factor(
      domain,
      "BOD"= "Bodily sensation",
      "NEG" = "Negative affect",
      "POS" = "Social connection",
      "COG" = "Cognition and control"),
    dev_factor = recode_factor(
      response,
      "The child is biologically \"preprogrammed\" to have this ability" = "preprogrammed",
      "The child has experiences in the womb that give them this ability" = "womb_experiences",
      "The child's body grows and matures (for example, muscles get stronger, child gets taller)" = "body_grows",
      "The child's brain changes (for example, brain grows bigger, more or fewer connections between neurons)" = "brain_changes",
      "The child's senses improve (for example, vision gets sharper, hearing improves)" = "senses_improve",
      "The child observes the objects and the physical world around him or her" = "observes_objects",
      "The child observes the people around him or her" = "observes_people",
      "The child interacts with the people around him or her" = "interacts_people",
      "People explicitly teach the child how to do this" = "people_teach",
      "The child actively experiments with how to do this" = "experiments",
      .default = "other"),
    dev_factor = relevel(dev_factor, ref = "other"),
    dev_factor_text = recode_factor(
      dev_factor,
      "preprogrammed" = "biologically 'preprogrammed'",
      "womb_experiences" = "experiences in the womb",
      "body_grows" = "body grows and matures",
      "brain_changes" = "brain changes",
      "senses_improve" = "senses improve",
      "observes_objects" = "observes objects and physical world",
      "experiments" = "actively experiments",
      "observes_people" = "observes people",
      "interacts_people" = "interacts with people",
      "people_teach" = "people explicitly teach"),
    capacity = factor(
      capacity,
      levels = c(
        "feeling_pain", "getting_hungry", 
        "feeling_distressed", "feeling_helpless",
        "feeling_happy", "learning_from_other_people",
        "controlling_their_emotions", "reasoning_about_things"))) %>%
  select(-domain, -response)

# contrasts(d3_mech_choice$factor) <- cbind(neg_bod = c(0, 1, 0, 0),
#                                           soc_bod = c(0, 0, 1, 0),
#                                           cog_bod = c(0, 0, 0, 1))

contrasts(d3_mech_choice$factor) <- cbind(bod_gm = c(1, -1, 0, 0),
                                          soc_gm = c(0, -1, 1, 0),
                                          cog_gm = c(0, -1, 0, 1))

# contrasts(d3_mech_choice$factor) <- cbind(bod_gm = c(1, 0, -1, 0),
#                                            neg_gm = c(0, 1, -1, 0),
#                                            cog_gm = c(0, 0, -1, 1))

contrasts(d3_mech_choice$dev_factor) <- cbind( # dummy = senses_improve
  preprogrammed =     c(rep(0, 0), 1, rep(0, 3), -1, rep(0, 6)),
  womb_experiences =  c(rep(0, 1), 1, rep(0, 2), -1, rep(0, 6)),
  body_grows =        c(rep(0, 2), 1, rep(0, 1), -1, rep(0, 6)),
  brain_changes =     c(rep(0, 3), 1, rep(0, 0), -1, rep(0, 6)),
  observes_objects =  c(rep(0, 4), -1, rep(0, 0), 1, rep(0, 5)),
  observes_people =   c(rep(0, 4), -1, rep(0, 1), 1, rep(0, 4)),
  interacts_people =  c(rep(0, 4), -1, rep(0, 2), 1, rep(0, 3)),
  people_teach =      c(rep(0, 4), -1, rep(0, 3), 1, rep(0, 2)),
  experiments =       c(rep(0, 4), -1, rep(0, 4), 1, rep(0, 1)),
  other =             c(rep(0, 4), -1, rep(0, 5), 1, rep(0, 0)))

# contrasts(d3_mech_choice$dev_factor) <- cbind( # dummy = other
#   preprogrammed =     c(rep(0, 0), 1, rep(0, 9), -1),
#   womb_experiences =  c(rep(0, 1), 1, rep(0, 8), -1),
#   body_grows =        c(rep(0, 2), 1, rep(0, 7), -1),
#   brain_changes =     c(rep(0, 3), 1, rep(0, 6), -1),
#   senses_improve =    c(rep(0, 4), 1, rep(0, 5), -1),
#   observes_objects =  c(rep(0, 5), 1, rep(0, 4), -1),
#   observes_people =   c(rep(0, 6), 1, rep(0, 3), -1),
#   interacts_people =  c(rep(0, 7), 1, rep(0, 2), -1),
#   people_teach =      c(rep(0, 8), 1, rep(0, 1), -1),
#   experiments =       c(rep(0, 9), 1, rep(0, 0), -1))
```

#### Summary tables

```{r s3 mech choice summary stats}
# overall
d3_mech_choice_sum_a <- d3_mech_choice %>%
  count(dev_factor_text) %>%
  complete(dev_factor_text, fill = list(n = 0)) %>%
  mutate(prop = n/sum(n))

d3_mech_choice_sum_a %>%
  mutate(percent = scales::percent(n/sum(n), accuracy = 1)) %>%
  select(-prop) %>%
  arrange(dev_factor_text) %>%
  rename(`Developmental mechanism` = dev_factor_text,
         `%` = percent) %>%
  kbl() %>%
  kable_styling() %>%
  collapse_rows(1, valign = "top")

# by domain
d3_mech_choice_sum_b <- d3_mech_choice %>%
  count(factor, dev_factor_text) %>%
  complete(dev_factor_text, nesting(factor), fill = list(n = 0)) %>%
  group_by(factor) %>%
  mutate(prop = n/sum(n)) %>%
  ungroup() %>%
  select(factor, everything())

d3_mech_choice_sum_b %>%
  mutate(percent = scales::percent(n/sum(n), accuracy = 1)) %>%
  select(-prop) %>%
  arrange(factor, dev_factor_text) %>%
  rename(Domain = factor,
         `Developmental mechanism` = dev_factor_text,
         `%` = percent) %>%
  kbl() %>%
  kable_styling() %>%
  collapse_rows(1, valign = "top")

# by domain and capacity
d3_mech_choice_sum_c <- d3_mech_choice %>%
  count(factor, capacity, dev_factor_text) %>%
  complete(dev_factor_text, nesting(factor, capacity), fill = list(n = 0)) %>%
  group_by(factor, capacity) %>%
  mutate(prop = n/sum(n)) %>%
  ungroup() %>%
  select(factor, capacity, everything())

d3_mech_choice_sum_c %>%
  mutate(percent = scales::percent(n/sum(n), accuracy = 1)) %>%
  select(-prop) %>%
  arrange(factor, capacity, dev_factor_text) %>%
  mutate(capacity = gsub("_", " ", capacity)) %>%
  rename(Domain = factor,
         Capacity = capacity,
         `Developmental mechanism` = dev_factor_text,
         `%` = percent) %>%
  kbl() %>%
  kable_styling() %>%
  collapse_rows(1:2, valign = "top")
```

```{r d3 mech choice modes}
d3_mech_choice_modes_strict <- d3_mech_choice_sum_c %>% 
  left_join(d3_mech_choice %>% distinct(dev_factor, dev_factor_text)) %>%
  select(-n, -dev_factor_text) %>%
  group_by(factor, capacity) %>%
  slice_max(n = 1, order_by = prop, with_ties = T) %>%
  ungroup() %>%
  rename(modal_dev_factor = dev_factor,
         modal_prop = prop) %>%
  select(factor, capacity, modal_dev_factor, modal_prop)

d3_mech_choice_modes_within10 <- d3_mech_choice_sum_c %>% 
  left_join(d3_mech_choice %>% distinct(dev_factor, dev_factor_text)) %>%
  select(-dev_factor_text) %>%
  select(dev_factor, factor, capacity, prop) %>%
  left_join(d3_mech_choice_modes_strict) %>%
  mutate(prop_diff = modal_prop - prop) %>%
  filter(abs(prop_diff) <= 0.10) %>%
  select(factor, capacity, dev_factor, prop, everything())
```

#### Multinomial regression

Here we present a multinomial log-linear model to predict the likelihood of selecting a mechanism by capacity Using the "nnet" package for R [@venables.ripley2002], we specified this model as follows: _multinom(choice of extrinsic mechanism ~ capacity_. (Note that this model was not equipped to represent the within-subjects nature of this dataset, and that capacities are not nested within domain for these purposes.) In this analysis, capacity was effect-coded for comparisons to the grand mean collapsing across capacities. The full results of this model are presented in Supplementary Table \@ref(tab:tabs3multinom).

This analysis aligned with our observation that participants' choice of mechanisms as the "most important" driver of development varied by capacity.

```{r multinom}
# effect coding, drop feeling_helpless
contrasts(d3_mech_choice$capacity) <- cbind(
  feeling_pain =                c(rep(0, 0),  1, rep(0, 2), -1, rep(0, 4)),
  getting_hungry =              c(rep(0, 1),  1, rep(0, 1), -1, rep(0, 4)),
  feeling_distressed =          c(rep(0, 2),  1, rep(0, 0), -1, rep(0, 4)),
  feeling_happy =               c(rep(0, 3), -1, rep(0, 0),  1, rep(0, 3)),
  learning_from_other_people =  c(rep(0, 3), -1, rep(0, 1),  1, rep(0, 2)),
  controlling_their_emotions =  c(rep(0, 3), -1, rep(0, 2),  1, rep(0, 1)),
  reasoning_about_things =      c(rep(0, 3), -1, rep(0, 3),  1, rep(0, 0))
)

s3_multinom <- nnet::multinom(
  dev_factor ~ capacity, 
  data = d3_mech_choice)

# summary(s3_multinom)
# parameters(s3_multinom, standardize = "refit")
```

```{r tabs3multinom, results = "asis", include = T, message = F, warning = F}
cat("\\begin{landscape}")
parameters(s3_multinom, standardize = "refit") %>%
  select(Response, Parameter, Coefficient, SE, CI_low, CI_high, z, p) %>%
  rename(dev_factor = Response) %>%
  left_join(d3_mech_choice %>% distinct(dev_factor, dev_factor_text)) %>%
  select(-dev_factor) %>%
  rename(Response = dev_factor_text) %>%
  mutate(Parameter = case_when(
    Parameter == "(Intercept)" ~ "Grand mean (GM) (intercept)",
    TRUE ~ paste0("Capacity (", str_to_sentence(gsub("_", " ", gsub("capacity", "", Parameter))), " vs. GM)"))) %>%
  mutate(p = ifelse(p < 0.001, "<0.001", round_x(p, 3))) %>%
  mutate(`95% CI` = paste0("[", round_x(CI_low), ", ", round_x(CI_high), "]")) %>%
  select(-starts_with("CI")) %>%
  select(Response, Parameter, Coefficient, `95% CI`, everything()) %>%
  kable("latex", booktabs = T, longtable = T,
        # caption = "\\text{Study 3, multinomial model results.}",
        caption = "Study 3, multinomial log linear model results.",
        digits = 2, align = c("lrrrrrrrrrrrrrr")) %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8) %>%
  collapse_rows(1, valign = "top")
cat("\\end{landscape}")
```

```{r multinom signif, include = F}
# take a closer look at only those effects where p<0.01
s3_multinom_signif <- parameters(s3_multinom, standardize = "refit") %>%
  data.frame() %>% 
  filter(p < 0.01) %>%
  filter(!grepl("Intercept", Parameter)) %>%
  mutate(Parameter = gsub("capacity", "", Parameter))

s3_multinom_signif %>%
  select(Parameter, Response, Coefficient) %>%
  rename(capacity = Parameter,
         dev_factor = Response) %>%
  left_join(d3_cap_ratings %>% distinct(factor, capacity)) %>%
  left_join(d3_mech_choice %>% distinct(dev_factor, dev_factor_text)) %>%
  mutate(direction = ifelse(Coefficient > 0, "positive", "negative")) %>%
  select(factor, direction, capacity, dev_factor_text, everything()) %>%
  select(-dev_factor) %>%
  arrange(factor, direction, capacity, desc(abs(Coefficient))) %>%
  mutate(capacity = gsub("_", " ", capacity)) %>%
  rename(Domain = factor,
         Direction = direction,
         Capacity = capacity,
         `Develompental mechanism` = dev_factor_text) %>%
  kbl(digits = 2) %>%
  kable_styling() %>%
  collapse_rows(1:3, valign = "top")
```

In brief, this analysis suggested that selections of what we would call "social" mechanisms of development (explicit teaching and interactions with people) were particularly uncommon for capacities in the domain of _bodily sensations_, and particularly common for capacities in the domain of _cognition and control_ (as well as one of the capacities in the domain of _social connection_). More broadly, many of the developmental mechanisms were selected more often for capacities in the domain of _cognition and control_ than capacities in other domains, reflecting the greater diversity of choices in this domain, including the (occasional) selection of mechanisms that were rarely if ever selected as 'most important' in other domains (e.g., active experimentation on the part of the child). 

```{r fig03, fig.cap="Perceived importance of various mechanisms in the development of four domains of mental life among US adults (Study 3); see main text for the full text of each mechanism. Small, lighter points represent individual participants’ ratings of how important each developmental mechanism is in the development of each capacity within a domain (first row: bodily sensation; second row: negative affect; third row: social connection; fourth row: cognition and control). Larger black and red points correspond to mean importance ratings across the sample, and error bars are bootstrapped 95% confidence intervals. On the right side of each panel are the percentages of participants who selected each developmental mechanism as the ‘most important’ driver of development for that capacity, with modal selections in red.", fig.width = 10, fig.asp = 1.2}
d3_mech_ratings %>%
  left_join(d3_mech_choice_sum_c %>% select(dev_factor_text, factor, capacity, prop)) %>%
  left_join(d3_mech_choice_modes_within10 %>% mutate(mode = T)) %>%
  mutate(mode = ifelse(is.na(mode), F, T),
         capacity = str_wrap(gsub("_", " ", capacity), 100)) %>% 
  ggplot(aes(x = reorder(dev_factor_text, desc(as.numeric(dev_factor_text))), 
             y = response)) +
  facet_wrap(facets = vars(reorder(capacity, as.numeric(factor))), ncol = 2,
             scales = "free_x") + # to repeat axis labels
  ggnewscale::new_scale_color() +
  geom_jitter(
    aes(color = factor), 
    height = 0.2, size = 0.1, alpha = 0.2,
    show.legend = F) +
  scale_color_viridis_d(
    name = "Domain:",
    end = 0.9, option = "viridis") +
  ggnewscale::new_scale_color() +
  geom_pointrange( # 95% CIs
    data = . %>%
      group_by(factor, capacity, dev_factor_text, mode) %>%
      multi_boot_standard(col = "response", na.rm = T) %>%
      ungroup(),
    aes(y = mean, ymin = ci_lower, ymax = ci_upper, color = mode)) +
  # geom_pointrange( # standard deviations
  #   data = . %>%
  #     group_by(factor, capacity, dev_factor_text) %>%
  #     summarise(mean = mean(response, na.rm = T),
  #               sd = sd(response, na.rm = T),
  #               n = n(),
  #               ci_lower = mean - sd,
  #               ci_upper = mean + sd) %>%
  #     ungroup(),
  #   aes(y = mean, ymin = ci_lower, ymax = ci_upper)) +
  scale_color_manual(
    name = "Modal choice of 'most important' developmental mechanism:", 
    values = c("black", "#f94802")) +
  geom_text(
    data = . %>%
      distinct(factor, capacity, dev_factor, dev_factor_text, prop, mode),
    aes(y = 6.5, label = paste0(round(prop, 2)*100, "%"),
        color = mode),
    size = 2, hjust = 0,
    show.legend = F) +
  annotate(geom = "text", label = "'most important'\nmechanism", 
           size = 2, hjust = 0.5, vjust = 0, 
           x = 11, y = 6.5) +
  theme_minimal() + 
  theme(panel.border = element_rect(fill = NA, linewidth = 0.2),
        panel.spacing.x = unit(2, "lines"),
        legend.position = "bottom",
        legend.box = "vertical",
        legend.spacing = unit(0, unit = "lines"),
        legend.box.spacing = unit(0, unit = "lines"),
        plot.margin = unit(c(1,3,1,1), "lines")) +
  scale_y_continuous(limits = c(NA, NA), breaks = seq(0, 6, 1)) +
  coord_flip(ylim = c(0, 6), xlim = c(1, 10), clip = "off") +
  labs(y = "\nImportance rating (0 = not at all important, 6 = extremely important)", 
       x = "Developmental mechanism")

# save plot
if (resave_plots == T | !file.exists("../outputs/fig03.pdf")) {
  ggsave("../outputs/fig03.svg")
  ggsave("../outputs/fig03.pdf")
}
```

```{r fig03_jpg, fig.cap="Perceived importance of various mechanisms in the development of four domains of mental life among US adults (Study 3); see main text for the full text of each mechanism. Small, lighter points represent individual participants’ ratings of how important each developmental mechanism is in the development of each capacity within a domain (first row: bodily sensation; second row: negative affect; third row: social connection; fourth row: cognition and control). Larger black and red points correspond to mean importance ratings across the sample, and error bars are bootstrapped 95% confidence intervals. On the right side of each panel are the percentages of participants who selected each developmental mechanism as the ‘most important’ driver of development for that capacity, with modal selections in red.", fig.width = 10, fig.asp = 1.2}
d3_mech_ratings %>%
  left_join(d3_mech_choice_sum_c %>% select(dev_factor_text, factor, capacity, prop)) %>%
  left_join(d3_mech_choice_modes_within10 %>% mutate(mode = T)) %>%
  mutate(mode = ifelse(is.na(mode), F, T),
         capacity = str_wrap(gsub("_", " ", capacity), 100)) %>% 
  ggplot(aes(x = reorder(dev_factor_text, desc(as.numeric(dev_factor_text))), 
             y = response)) +
  facet_wrap(facets = vars(reorder(capacity, as.numeric(factor))), ncol = 2,
             scales = "free_x") + # to repeat axis labels
  ggnewscale::new_scale_color() +
  geom_jitter(
    aes(color = factor), 
    height = 0.2, size = 0.1, alpha = 0.4,
    show.legend = F) +
  scale_color_viridis_d(
    name = "Domain:",
    end = 0.9, option = "viridis") +
  ggnewscale::new_scale_color() +
  geom_pointrange( # 95% CIs
    data = . %>%
      group_by(factor, capacity, dev_factor_text, mode) %>%
      multi_boot_standard(col = "response", na.rm = T) %>%
      ungroup(),
    aes(y = mean, ymin = ci_lower, ymax = ci_upper, color = mode)) +
  # geom_pointrange( # standard deviations
  #   data = . %>%
  #     group_by(factor, capacity, dev_factor_text) %>%
  #     summarise(mean = mean(response, na.rm = T),
  #               sd = sd(response, na.rm = T),
  #               n = n(),
  #               ci_lower = mean - sd,
  #               ci_upper = mean + sd) %>%
  #     ungroup(),
  #   aes(y = mean, ymin = ci_lower, ymax = ci_upper)) +
  scale_color_manual(
    name = "Modal choice of 'most important' developmental mechanism:", 
    values = c("black", "#f94802")) +
  geom_text(
    data = . %>%
      distinct(factor, capacity, dev_factor, dev_factor_text, prop, mode),
    aes(y = 6.5, label = paste0(round(prop, 2)*100, "%"),
        color = mode),
    size = 2, hjust = 0,
    show.legend = F) +
  annotate(geom = "text", label = "'most important'\nmechanism", 
           size = 2, hjust = 0.5, vjust = 0, 
           x = 11, y = 6.5) +
  theme_minimal() + 
  theme(panel.border = element_rect(fill = NA, linewidth = 0.2),
        panel.spacing.x = unit(2, "lines"),
        legend.position = "bottom",
        legend.box = "vertical",
        legend.spacing = unit(0, unit = "lines"),
        legend.box.spacing = unit(0, unit = "lines"),
        plot.margin = unit(c(1,3,1,1), "lines")) +
  scale_y_continuous(limits = c(NA, NA), breaks = seq(0, 6, 1)) +
  coord_flip(ylim = c(0, 6), xlim = c(1, 10), clip = "off") +
  labs(y = "\nImportance rating (0 = not at all important, 6 = extremely important)", 
       x = "Developmental mechanism")

# save plot
if (resave_plots == T | !file.exists("../outputs/fig03.jpg")) {
  ggsave("../outputs/fig03.jpg")
}
```

```{r s3 export results}
# save numerical results
s3_results <- list(demo = d3_demo,
                   cap_rat_ss = d3_cap_ratings_sum,
                   mech_rat_ss_a = d3_mech_ratings_sum_a,
                   mech_rat_ss_b = d3_mech_ratings_sum_b,
                   mech_rat_ss_c = d3_mech_ratings_sum_c,
                   mech_rat_ss_d = d3_mech_ratings_sum_d,
                   gam = s3_gam, 
                   lmer = s3_lmer, 
                   lmer_bod = s3_lmer_bod,
                   lmer_neg = s3_lmer_neg,
                   lmer_soc = s3_lmer_soc,
                   lmer_cog = s3_lmer_cog,
                   mech_choice_ss_a = d3_mech_choice_sum_a,
                   mech_choice_ss_b = d3_mech_choice_sum_b,
                   mech_choice_ss_c = d3_mech_choice_sum_c,
                   mech_choice_modes = d3_mech_choice_modes_within10) 

saveRDS(s3_results, "../outputs/s3_results.rds")
```


# Study comparison: Capacity attributions to newborns, 9-months, and 5-year-olds

As described in the main text, all of the current studied employed a fixed-order design, in which all participants assessed target ages in a fixed order beginning with the youngest target age (newborns) and proceeding chronologically until the oldest target age (5-year-olds). This decision reflects the fact that our priority in these studies was to assess beliefs about the developmental trajectories of capacities, rather than beliefs about capacities at a specific point in development. However, one limitation of this fixed-order design is that it makes it difficult to assess the stability of participants’ assessments of the fine-grained differences between specific points in development. 

To provide some sense of the stability and reliability of US adults' representations of specific target ages across studies, Supplementary Figure \@ref(fig:figS06) presents a comparison of capacity ratings and summary scores by domain at the three target ages that were included in all of the current studies (newborns, 9-month-olds, and 5-year-olds). As a reminder, in Study 1, participants assessed 60 capacities (including but not limited to the 20 capacities depicted in Panel A), at each of the 3 target ages depicted here. In Study 2, participants assessed the 20 capacities depicted in Panel A, at 13 target ages (including but not limited to the 3 target ages depicted here). In Study 3, participants assessed 8 capacities (a subset of the 20 capacities depicted in Panel A), at 13 target ages (including but not limited to the 3 target ages depicted here).

```{r}
d123_gam <- full_join(
  d1_gam %>% 
    mutate(study = "Study 1"),
  d2_gam %>% 
    mutate(study = "Study 2") %>%
    mutate(response = response*100)) %>%
  left_join(
    d3_gam %>% 
      distinct(target, target_ord)) %>%
  full_join(
    d3_gam %>% 
      mutate(study = "Study 3") %>%
      mutate(response = response*100))
```

```{r figS06 code, fig.width = 10, fig.asp = 1.2}
figS06 <- d123_gam %>%
  filter(target_ord %in% c("newborns", "9-month-olds", "5-year-olds")) %>%
  filter(capacity %in%
           c(full_join(d2_gam %>% distinct(capacity),
                       d3_gam %>% distinct(capacity)) %>%
               distinct(capacity) %>%
               pull(capacity) %>%
               as.character())) %>%
  complete(
    capacity, nesting(study, target, target_year, target_num, target_ord),
    fill = list(response = -1)) %>% # introduce -1 to align studies across panels
  select(-factor) %>%
  left_join(d2_gam %>% distinct(capacity, factor)) %>%
  left_join(d2_scales %>% select(-loading)) %>%
  mutate(capacity_text = str_wrap(gsub("_", " ", capacity), 40)) %>%
  ggplot(aes(x = target_ord, y = response, 
             color = factor)) +
  facet_wrap(facets = vars(reorder(capacity_text, as.numeric(order))), nrow = 4) +
  geom_point(
    aes(group = study),
    size = 0.02,
    alpha = 0.6,
    position = position_jitterdodge(
      dodge.width = 0.8,
      jitter.width = 0.6,
      jitter.height = 0)) +
  scale_color_viridis_d(end = 0.9, option = "viridis") +
  geom_path(
    data = . %>%
      group_by(study, target, target_year, target_num, target_ord, 
               factor, capacity, capacity_text, order) %>%
      summarise(mean = mean(response, na.rm = T)) %>%
      ungroup(),
    aes(y = mean, group = study),
    color = "black", lty = 2, 
    position = position_dodge(width = 0.8)) +
  geom_pointrange(
    data = . %>%
      group_by(study, target, target_year, target_num, target_ord, 
               factor, capacity, capacity_text, order) %>%
      multi_boot_standard(col = "response") %>%
      ungroup(),
    aes(y = mean, ymin = ci_lower, ymax = ci_upper, 
        group = study, fill = study),
    color = "black", shape = 21, fatten = 6,
    position = position_dodge(width = 0.8)) +
  geom_text(
    data = . %>%
      group_by(study, target, target_year, target_num, target_ord, 
               factor, capacity, capacity_text, order) %>%
      summarise(mean = mean(response, na.rm = T)) %>%
      ungroup(),
    aes(y = mean, group = study, label = gsub("Study ", "", study)),
    color = "white", size = 2.5, fontface = "bold", hjust = 0.5, vjust = 0.5,
    position = position_dodge(width = 0.8)) +
  scale_fill_manual(values = c("#9801a3", "#a3012f", "#a36501")) +
  scale_y_continuous(
    limits = c(-0.05, NA), # exclude the specious "-1" responses introduced above
    breaks = seq(0, 100, 25),
    labels = seq(0, 100, 25)) +
  scale_x_discrete(labels = c("newborns", "9 months", "5 years")) +
  theme_minimal() +
  theme(panel.background = element_rect(fill = NA), 
        strip.text = element_text(vjust = 0),
        legend.position = "top") +
        # axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
        # plot.margin = unit(c(0.5, 0.5, 0.5, 1.1), "cm")) +
  labs(x = "Target age",
       y = "Mean response\n0 = not at all capable, 100 = completely capable",
       color = "Domain:",
       fill = "Study:") +
  guides(color = guide_legend(override.aes = list(alpha = 1, size = 2)))
figS06

if (resave_plots == T | !file.exists("../outputs/figS06.pdf")) {
  ggsave("../outputs/figS06.svg")
  ggsave("../outputs/figS06.pdf")
}
```

```{r figS06, fig.cap="Comparison of capacity ratings across studies (Studies 1-3). This visualization includes responses for each of the capacities and target ages held in common across at least two of our three studies. Lighter colorful represent individual participants’ responses or scores, larger points correspond to mean responses or scores across the sample, and error bars are bootstrapped 95% confidence intervals.", include = T}
knitr::include_graphics("../outputs/figS06.pdf")
```


# Selection of capacities: Comparison to previous studies

Supplementary Table \@ref(tab:tabcapcomp) provides an item-by-item comparison of the capacities used here to those used in previous studies of mind perception [@gray.etal2007; @weisman.etal2017a; @weisman.etal2017b; @malle2019].

```{r capacities comparison df}
d_cap_comp <- read_csv("../data/capacities_comparison.csv") %>%
  mutate(across(where(is.character), ~tolower(.))) %>%
  mutate(across(where(is.character), ~gsub("recognizing others.", "recognizing others'", .))) %>%
  mutate(across(where(is.character), ~gsub("self.control", "self-control", .))) %>%
  left_join(d1_scales %>%
              mutate(capacity = gsub("_", " ", capacity),
                     capacity = gsub("recognizing others ", "recognizing others' ", capacity),
                     capacity = gsub("self.control", "self-control", capacity)) %>%
              rename(`Study 1` = capacity)) %>%
  rename(`Weisman et al. (2017, PNAS)` = `Weisman et al. (2017a)`,
         `Weisman et al. (2017, Proc. CogSci)` = `Weisman et al. (2017b)`) %>%
  mutate(across(c(`Weisman et al. (2017, PNAS)`, `Weisman et al. (2017, Proc. CogSci)`, 
                  `Gray et al. (2007)`, `Malle (2019, Study 1)`, `Malle (2019, Study 2)`),
                ~case_when(is.na(`Study 1`) & grepl("similar to", .) ~ NA_character_, TRUE ~ .))) %>%
  arrange(order, `Weisman et al. (2017, PNAS)`, `Weisman et al. (2017, Proc. CogSci)`, 
          `Gray et al. (2007)`, `Malle (2019, Study 1)`, `Malle (2019, Study 2)`) %>%
  filter(!(is.na(`Study 1`) & 
             is.na(`Weisman et al. (2017, PNAS)`) &
             is.na(`Weisman et al. (2017, Proc. CogSci)`) &
             is.na(`Gray et al. (2007)`) &
             is.na(`Malle (2019, Study 1)`) & 
             is.na(`Malle (2019, Study 2)`))) %>%
  mutate(across(where(is.character), ~replace_na(., "-"))) %>%
  mutate(factor = case_when(is.na(factor) ~ "Not included", TRUE ~ as.character(factor)),
         factor = factor(factor, 
                         levels = c("Bodily sensation", "Negative affect",
                                    "Social connection", "Cognition and control", 
                                    "Not included"))) %>%
  select(factor, everything(), -loading, -order) %>%
  rename(Domain = factor) %>%
  ungroup()
```

```{r tabcapcomp, results = "asis", include = T, warning = F, message = F}
cat("\\begin{landscape}")
tabcapcomp <- d_cap_comp %>%
  select(-Domain) %>%
  mutate_all(~gsub("similar to .", "similar to \\`", .)) %>% 
  mutate(`Study 1` = case_when(
    `Study 2` != "-" ~ paste0(`Study 1`, footnote_marker_symbol(1)),
    TRUE ~ `Study 1`)) %>%
  mutate(`Study 1` = case_when(
    `Study 3` != "-" ~ paste0(`Study 1`, footnote_marker_symbol(2)),
    TRUE ~ `Study 1`)) %>%
  select(-`Study 2`, -`Study 3`) %>%
  mutate(across(c(`Gray et al. (2007)`),
                ~gsub("\\..*", "", .))) %>%
  kable("latex", booktabs = T, escape = F, longtable = T,
        caption = "Item-by-item comparison of the capacities included in the current studies to capacities included in previous studies of mind perception.") %>%
  kable_styling(latex_options = c("repeat_header"), font_size = 8) %>%
  add_header_above(c("Current studies" = 1,
                     "Previous studies" = 5)) %>%
  # column_spec(1, width = "4cm") %>%
  column_spec(2:6, width = "2.7cm") %>%
  pack_rows(group_label= "Bodily sensation", start_row = 1, end_row = 11) %>%
  pack_rows("Negative affect", 12, 16) %>%
  pack_rows("Social connection", 17, 38) %>%
  pack_rows("Cognition and control", 39, 60) %>%
  pack_rows("Not included", 61, nrow(d_cap_comp)) %>%
  footnote(symbol = c("Also used in Study 2; ", "Also used in Study 3"),
           symbol_title = "Notes: ",
           footnote_as_chunk = T)
tabcapcomp
cat("\\end{landscape}")
```


\newpage
# References
